### 2017-07-30 rebuttal2.txt

The author(s) thank the three reviewers for their feedback, which is extremely useful for a revised version of the manuscript, even in case of non-acceptance. Here are our comments point-by-point.

Reviewer 1 ("Assigned_Reviewer_1")

1. The "sample level" maximum-entropy method has been repeatedly used and discussed in computational neuroscience since the work of Bohte et al (2000). A more complete bibliography is given at the end of this response and could be easily extended. Unfortunately our introduction did not clarify and reference this context enough; it would surely be amended in a revised version. The method is also used by non-experts; our purpose is to clarify some implicit assumptions that experts and non-experts seem to be unaware of.

2. Our discussion starts from an observation: in the neuroscientific literature some authors generalize conclusions drawn via maximum-entropy on a sample of recorded neurons, to whole areas the sample is recorded from, eg retina or motor cortex. Unfortunately this observation was not made clear in our introduction. Our purpose is to give a warning: the typical sample-level approach (Eq. 17) is incompatible with the beyond-sample generalizations made by some authors. We propose a simple alternative consistent with this assumption (Eqs 15, 19).

The case of absolute ignorance (I'') is not used in the rest of the manuscript; we apologize for the confusion it creates. We would omit it in a revised version.

We invite the reviewer to read also our responses 1 & 2 to the second reviewer.

3. In the manuscript the principal equations are:

(15) to be used when a larger population is assumed and its size roughly known;

(17) to be used when no larger population is assumed: most literature uses this equation but also assumes a larger population;

(19) to be used when the the size of the larger population is uncertain.

Equations (15) and (19) are novel, to our knowledge. Unfortunately we did not emphasize the role of these three equations and give clear guidance to their use. This should be amended in a revised version.

4. The captions should indeed give a clearer explanation of the figures. Owing to the page limit we preferred not to repeat explanations given in the main text.

Reviewer 2 ("Assigned_Reviewer_2")

1. The starting points of the manuscripts are three:

- the *sample-level* maximum-entropy method is increasingly being used in the neuroscientific literature (references below);

- this literature often generalizes results obtained with this method from a recorded sample to a whole area;

- worries have been expressed (eg Ganmor et al 2011, Roudi et al PlosCB 2009) about an excessive sample-size dependence of maximum-entropy results.

Unfortunately these three starting points are not clarified enough in our introduction; this should be amended in a revised version. We motivate the sample-size worries by simple probabilistic considerations, and provide alternative formulae to be used: Eq. (15) when the size of the full population is known, Eq. (19) when it is uncertain.

2. The sample approach is routinely used in the literature and we agree with the reviewer: "it is not even an option" if the recorded neurons are assumed to be a sample of a larger population. This is the message of the manuscript, unfortunately not clear enough. Its analysis is meant to explain why the sample approach is not an option and to propose two corrected formulae (Eqs 15, 19).

Our phrase "left for future work" was unfortunate: Eq. (19) for uncertain population size is actually an end result, ready to be used computationally if not analytically. What we left for future work was the analysis of its mathematical difference from (15) and (17). In a revised version of the paper we would add an example of distribution (19) to Figs 2 & 3.

We invite the reviewer to read also our responses 1, 2, 3 to the first reviewer.

3. Large space was dedicated to deriving the formulae of sampling without replacement. As mentioned on lines 42-45 this was on purpose because these old formulae are central in many neuroscientific problems and yet we have rarely seen them used in the neuroscience literature. We wanted to stress their importance.

But from those formulae the manuscript also presents novel results: Eqs (15) & (19), which can immediately be used instead of Eq. (17), currently used in the literature.

4. Equation (12) and its discussion (lines 113-129) are an excursion irrelevant to the rest of the paper, to the discussion of maximum-entropy and to the main equations we derive. We would eliminate this whole passage in a revised version.

5. We would have liked to present a Bayesian analysis of the problem, but the NIPS page limit prevented this. A measure of discrepancy, eg Kullback-Leibler divergence, between Eqs (15), (17), (19) could be added in a revised version.

Reviewer 3 ("Assigned_Reviewer_5")

We cordially thank the reviewer for the appreciative words. We want to add that our figures *do* use actual recorded neural activity data, but because of the double-blind review procedure we omitted this information, which we plan to add in the final version if the manuscript were accepted.

Sample references of maximum-entropy use in neuroscientific literature:
Martignon & al, Biol Cybern (1995)
Bohte & al, Neural Comp (2000)
Schneidman & al,  Nature (2006)
Shlens & al,  J Neurosci (2006)
Tkačik & al, arXiv:q-bio/0611072 (2006)
Macke & al, Neural Comp (2009)
Roudi & al, PLoS CB (2009)
Roudi & al, Front Comp Neurosci (2009)
Roudi & al, Phys Rev E (2009)
Tkačik & al, arXiv:09125409 (2009)
Macke & al, Tech Rep (2009)
Gerwinn & al, Front Comput Neurosci (2010)
Barreiro & al, arXiv:10112797 (2010)
Macke & al, NIPS (2011)
Macke & al, Phys Rev Lett (2011)
Ganmor & al, Proc Natl Acad Sci (2011)
Granot-Atedgi & al, PLoS CB (2013)
Macke & al, Entropy (2013)
Tkačik & al, Proc Natl Acad Sci (2014)
Mora & al, Phys Rev Lett (2015)
Shimazaki & al,  Sci Rep (2015)