### 2018-05-15 maxNt_v3.tex
V3

%% Author: PGL  Porta Mana
%% Created: 2015-11-04T09:20:08+0100
%% Last-Updated: 2018-05-15T22:50:14+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Report-no: ***
\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
% \PassOptionsToPackage{numbers, compress}{natbib}
% before loading nips_2017
%
% to avoid loading the natbib package, add option nonatbib:
% \usepackage[nonatbib]{nips_2017}

\usepackage[final,%
nonatbib]{nips_2017}

\newcommand*{\firstdraft}{4 November 2015}

\newcommand*{\pdftitle}{Representative samples\\ and maximum-entropy distributions:\\ a dilemma}

\newcommand*{\headtitle}{\pdftitle} \newcommand*{\pdfauthor}{P.G.L. Porta Mana, V. Rostami, E. Torre}
\usepackage{newtxmath}
%\usepackage{pifont}
%\usepackage{fontawesome}
\usepackage[T1]{fontenc} 
\input{glyphtounicode} \pdfgentounicode=1
\usepackage[utf8]{inputenx}
%\usepackage{newunicodechar}
% \newunicodechar{Ĕ}{\u{E}}
% \newunicodechar{ĕ}{\u{e}}
% \newunicodechar{Ĭ}{\u{I}}
% \newunicodechar{ĭ}{\u{\i}}
% \newunicodechar{Ŏ}{\u{O}}
% \newunicodechar{ŏ}{\u{o}}
% \newunicodechar{Ŭ}{\u{U}}
% \newunicodechar{ŭ}{\u{u}}
% \newunicodechar{Ā}{\=A}
% \newunicodechar{ā}{\=a}
% \newunicodechar{Ē}{\=E}
% \newunicodechar{ē}{\=e}
% \newunicodechar{Ī}{\=I}
% \newunicodechar{ī}{\={\i}}
% \newunicodechar{Ō}{\=O}
% \newunicodechar{ō}{\=o}
% \newunicodechar{Ū}{\=U}
% \newunicodechar{ū}{\=u}
% \newunicodechar{Ȳ}{\=Y}
% \newunicodechar{ȳ}{\=y}

\newcommand*{\bmmax}{3} % reduce number of bold fonts, before bm
\newcommand*{\hmmax}{0} % reduce number of heavy fonts, before bm
\usepackage{textcomp}
\usepackage{amsmath}
\usepackage{mathtools}
\setlength{\multlinegap}{0pt}

\usepackage{amssymb}
\usepackage{amsxtra}

\usepackage[british]{babel}\selectlanguage{british}
\newcommand*{\langfrench}{\foreignlanguage{french}}
\newcommand*{\langgerman}{\foreignlanguage{german}}
\newcommand*{\langitalian}{\foreignlanguage{italian}}
\newcommand*{\langswedish}{\foreignlanguage{swedish}}
\newcommand*{\langlatin}{\foreignlanguage{latin}}
\newcommand*{\langnohyph}{\foreignlanguage{nohyphenation}}

\usepackage[autostyle=false,autopunct=false,english=british]{csquotes}
\setquotestyle{british}

\let\openbox\relax
\usepackage{amsthm}
\newcommand*{\QED}{\textsc{q.e.d.}}
\renewcommand*{\qedsymbol}{\QED}
\theoremstyle{remark}
\newtheorem{note}{Note}
\newtheorem*{remark}{Note}
\newtheoremstyle{innote}{\parsep}{\parsep}{\footnotesize}{}{}{}{0pt}{}
\theoremstyle{innote}
\newtheorem*{innote}{}

\usepackage{bm}

\iffalse
\usepackage[shortlabels,inline]{enumitem}
\SetEnumitemKey{para}{itemindent=\parindent,leftmargin=0pt,listparindent=\parindent,parsep=0pt,itemsep=\topsep}
% \begin{asparaenum} = \begin{enumerate}[para]
% \begin{inparaenum} = \begin{enumerate*}
\setlist[enumerate,2]{label=\alph*.}
\setlist[enumerate]{leftmargin=\parindent}
\setlist[itemize]{leftmargin=\parindent}
\setlist[description]{leftmargin=\parindent}
\fi

%% With euler font cursive for Greek letters - the [1] means 100% scaling
\DeclareFontFamily{U}{egreek}{\skewchar\font'177}%
\DeclareFontShape{U}{egreek}{m}{n}{<-6>s*[0.95]eurm5 <6-8>s*[0.95]eurm7 <8->s*[0.95]eurm10}{}%
\DeclareFontShape{U}{egreek}{m}{it}{<->s*[0.95]eurmo10}{}%
\DeclareFontShape{U}{egreek}{b}{n}{<-6>s*[0.95]eurb5 <6-8>s*[0.95]eurb7 <8->s*[0.95]eurb10}{}%
\DeclareFontShape{U}{egreek}{b}{it}{<->s*[0.95]eurbo10}{}%
\DeclareSymbolFont{egreeki}{U}{egreek}{m}{it}%
\SetSymbolFont{egreeki}{bold}{U}{egreek}{b}{it}% from the amsfonts package
\DeclareSymbolFont{egreekr}{U}{egreek}{m}{n}%
\SetSymbolFont{egreekr}{bold}{U}{egreek}{b}{n}% from the amsfonts package
% Take also \sum, \prod, \coprod symbols from Euler fonts
\DeclareFontFamily{U}{egreekx}{\skewchar\font'177}
\DeclareFontShape{U}{egreekx}{m}{n}{%
       <-7.5>s*[0.9]euex7%
    <7.5-8.5>s*[0.9]euex8%
    <8.5-9.5>s*[0.9]euex9%
    <9.5->s*[0.9]euex10%
}{}
\DeclareSymbolFont{egreekx}{U}{egreekx}{m}{n}
\DeclareMathSymbol{\sumop}{\mathop}{egreekx}{"50}
\DeclareMathSymbol{\prodop}{\mathop}{egreekx}{"51}
\DeclareMathSymbol{\coprodop}{\mathop}{egreekx}{"60}
\makeatletter
\def\sum{\DOTSI\sumop\slimits@}
\def\prod{\DOTSI\prodop\slimits@}
\def\coprod{\DOTSI\coprodop\slimits@}
\makeatother
% Greek letters not usually given in LaTeX. Comment the unneeded ones
% \DeclareMathSymbol{\varpartial}{\mathalpha}{egreeki}{"40}
% \DeclareMathSymbol{\partialup}{\mathalpha}{egreekr}{"40}
% \DeclareMathSymbol{\alpha}{\mathalpha}{egreeki}{"0B}
% \DeclareMathSymbol{\beta}{\mathalpha}{egreeki}{"0C}
% \DeclareMathSymbol{\gamma}{\mathalpha}{egreeki}{"0D}
% \DeclareMathSymbol{\delta}{\mathalpha}{egreeki}{"0E}
% \DeclareMathSymbol{\epsilon}{\mathalpha}{egreeki}{"0F}
 \DeclareMathSymbol{\zeta}{\mathalpha}{egreeki}{"10}
% \DeclareMathSymbol{\eta}{\mathalpha}{egreeki}{"11}
% \DeclareMathSymbol{\theta}{\mathalpha}{egreeki}{"12}
% \DeclareMathSymbol{\iota}{\mathalpha}{egreeki}{"13}
 \DeclareMathSymbol{\kappa}{\mathalpha}{egreeki}{"14}
 \DeclareMathSymbol{\lambda}{\mathalpha}{egreeki}{"15}
% \DeclareMathSymbol{\mu}{\mathalpha}{egreeki}{"16}
 \DeclareMathSymbol{\nu}{\mathalpha}{egreeki}{"17}
 \DeclareMathSymbol{\xi}{\mathalpha}{egreeki}{"18}
% \DeclareMathSymbol{\omicron}{\mathalpha}{egreeki}{"6F}
% \DeclareMathSymbol{\pi}{\mathalpha}{egreeki}{"19}
% \DeclareMathSymbol{\rho}{\mathalpha}{egreeki}{"1A}
 \DeclareMathSymbol{\sigma}{\mathalpha}{egreeki}{"1B}
% \DeclareMathSymbol{\tau}{\mathalpha}{egreeki}{"1C}
% \DeclareMathSymbol{\upsilon}{\mathalpha}{egreeki}{"1D}
% \DeclareMathSymbol{\phi}{\mathalpha}{egreeki}{"1E}
% \DeclareMathSymbol{\chi}{\mathalpha}{egreeki}{"1F}
% \DeclareMathSymbol{\psi}{\mathalpha}{egreeki}{"20}
% \DeclareMathSymbol{\omega}{\mathalpha}{egreeki}{"21}
% \DeclareMathSymbol{\varepsilon}{\mathalpha}{egreeki}{"22}
% \DeclareMathSymbol{\vartheta}{\mathalpha}{egreeki}{"23}
% \DeclareMathSymbol{\varpi}{\mathalpha}{egreeki}{"24}
% \let\varrho\rho 
% \let\varsigma\sigma
% \let\varkappa\kappa
% \DeclareMathSymbol{\varphi}{\mathalpha}{egreeki}{"27}
% %
% \DeclareMathSymbol{\varAlpha}{\mathalpha}{egreeki}{"41}
% \DeclareMathSymbol{\varBeta}{\mathalpha}{egreeki}{"42}
% \DeclareMathSymbol{\varGamma}{\mathalpha}{egreeki}{"00}
% \DeclareMathSymbol{\varDelta}{\mathalpha}{egreeki}{"01}
% \DeclareMathSymbol{\varEpsilon}{\mathalpha}{egreeki}{"45}
 \DeclareMathSymbol{\varZeta}{\mathalpha}{egreeki}{"5A}
% \DeclareMathSymbol{\varEta}{\mathalpha}{egreeki}{"48}
% \DeclareMathSymbol{\varTheta}{\mathalpha}{egreeki}{"02}
 \DeclareMathSymbol{\varIota}{\mathalpha}{egreeki}{"49}
 \DeclareMathSymbol{\varKappa}{\mathalpha}{egreeki}{"4B}
 \DeclareMathSymbol{\varLambda}{\mathalpha}{egreeki}{"03}
% \DeclareMathSymbol{\varMu}{\mathalpha}{egreeki}{"4D}
% \DeclareMathSymbol{\varNu}{\mathalpha}{egreeki}{"4E}
 \DeclareMathSymbol{\varXi}{\mathalpha}{egreeki}{"04}
% \DeclareMathSymbol{\varOmicron}{\mathalpha}{egreeki}{"4F}
% \DeclareMathSymbol{\varPi}{\mathalpha}{egreeki}{"05}
% \DeclareMathSymbol{\varRho}{\mathalpha}{egreeki}{"50}
 \DeclareMathSymbol{\varSigma}{\mathalpha}{egreeki}{"06}
% \DeclareMathSymbol{\varTau}{\mathalpha}{egreeki}{"54}
 \DeclareMathSymbol{\varUpsilon}{\mathalpha}{egreeki}{"07}
% \DeclareMathSymbol{\varPhi}{\mathalpha}{egreeki}{"08}
% \DeclareMathSymbol{\varChi}{\mathalpha}{egreeki}{"58}
% \DeclareMathSymbol{\varPsi}{\mathalpha}{egreeki}{"09}
% \DeclareMathSymbol{\varOmega}{\mathalpha}{egreeki}{"0A} 
% %
% \DeclareMathSymbol{\Alpha}{\mathalpha}{egreekr}{"41}
% \DeclareMathSymbol{\Beta}{\mathalpha}{egreekr}{"42}
% \DeclareMathSymbol{\Gamma}{\mathalpha}{egreekr}{"00}
% \DeclareMathSymbol{\Delta}{\mathalpha}{egreekr}{"01}
% \DeclareMathSymbol{\Epsilon}{\mathalpha}{egreekr}{"45}
% \DeclareMathSymbol{\Zeta}{\mathalpha}{egreekr}{"5A}
% \DeclareMathSymbol{\Eta}{\mathalpha}{egreekr}{"48}
% \DeclareMathSymbol{\Theta}{\mathalpha}{egreekr}{"02}
% \DeclareMathSymbol{\Iota}{\mathalpha}{egreekr}{"49}
% \DeclareMathSymbol{\Kappa}{\mathalpha}{egreekr}{"4B}
% \DeclareMathSymbol{\Lambda}{\mathalpha}{egreekr}{"03}
% \DeclareMathSymbol{\Mu}{\mathalpha}{egreekr}{"4D}
% \DeclareMathSymbol{\Nu}{\mathalpha}{egreekr}{"4E}
% \DeclareMathSymbol{\Xi}{\mathalpha}{egreekr}{"04}
% \DeclareMathSymbol{\Omicron}{\mathalpha}{egreekr}{"4F}
% \DeclareMathSymbol{\Pi}{\mathalpha}{egreekr}{"05}
% \DeclareMathSymbol{\Rho}{\mathalpha}{egreekr}{"50}
% \DeclareMathSymbol{\Sigma}{\mathalpha}{egreekr}{"06}
% \DeclareMathSymbol{\Tau}{\mathalpha}{egreekr}{"54}
% \DeclareMathSymbol{\Upsilon}{\mathalpha}{egreekr}{"07}
% \DeclareMathSymbol{\Phi}{\mathalpha}{egreekr}{"08}
% \DeclareMathSymbol{\Chi}{\mathalpha}{egreekr}{"58}
% \DeclareMathSymbol{\Psi}{\mathalpha}{egreekr}{"09}
% \DeclareMathSymbol{\Omega}{\mathalpha}{egreekr}{"0A}
% %
% \DeclareMathSymbol{\alphaup}{\mathalpha}{egreekr}{"0B}
% \DeclareMathSymbol{\betaup}{\mathalpha}{egreekr}{"0C}
% \DeclareMathSymbol{\gammaup}{\mathalpha}{egreekr}{"0D}
% \DeclareMathSymbol{\deltaup}{\mathalpha}{egreekr}{"0E}
% \DeclareMathSymbol{\epsilonup}{\mathalpha}{egreekr}{"0F}
% \DeclareMathSymbol{\zetaup}{\mathalpha}{egreekr}{"10}
% \DeclareMathSymbol{\etaup}{\mathalpha}{egreekr}{"11}
% \DeclareMathSymbol{\thetaup}{\mathalpha}{egreekr}{"12}
% \DeclareMathSymbol{\iotaup}{\mathalpha}{egreekr}{"13}
% \DeclareMathSymbol{\kappaup}{\mathalpha}{egreekr}{"14}
% \DeclareMathSymbol{\lambdaup}{\mathalpha}{egreekr}{"15}
% \DeclareMathSymbol{\muup}{\mathalpha}{egreekr}{"16}
% \DeclareMathSymbol{\nuup}{\mathalpha}{egreekr}{"17}
% \DeclareMathSymbol{\xiup}{\mathalpha}{egreekr}{"18}
% \DeclareMathSymbol{\omicronup}{\mathalpha}{egreekr}{"6F}
%  \DeclareMathSymbol{\piup}{\mathalpha}{egreekr}{"19}
% \DeclareMathSymbol{\rhoup}{\mathalpha}{egreekr}{"1A}
% \DeclareMathSymbol{\sigmaup}{\mathalpha}{egreekr}{"1B}
% \DeclareMathSymbol{\tauup}{\mathalpha}{egreekr}{"1C}
% \DeclareMathSymbol{\upsilonup}{\mathalpha}{egreekr}{"1D}
% \DeclareMathSymbol{\phiup}{\mathalpha}{egreekr}{"1E}
% \DeclareMathSymbol{\chiup}{\mathalpha}{egreekr}{"1F}
% \DeclareMathSymbol{\psiup}{\mathalpha}{egreekr}{"20}
% \DeclareMathSymbol{\omegaup}{\mathalpha}{egreekr}{"21}
% \DeclareMathSymbol{\varepsilonup}{\mathalpha}{egreekr}{"22}
% \DeclareMathSymbol{\varthetaup}{\mathalpha}{egreekr}{"23}
% \DeclareMathSymbol{\varpiup}{\mathalpha}{egreekr}{"24}
% \let\varrhoup\rhoup 
% \let\varsigmaup\sigmaup
% \let\varkappaup\kappaup
% \DeclareMathSymbol{\varphiup}{\mathalpha}{egreekr}{"27}

%\newcommand*{\mathte}[1]{\textbf{\textit{\textsf{#1}}}}
% Upright sans-serif math alphabet
% \DeclareMathAlphabet{\mathsu}  {T1}{\sfdefault}{m}{n}
% \SetMathAlphabet{\mathsu}{bold}{T1}{\sfdefault}{b}{n}

\usepackage{mathdots}

\usepackage[usenames]{xcolor}
\definecolor{myblue}{RGB}{51,34,136}
\definecolor{mygreen}{RGB}{17,119,51}
\definecolor{myred}{RGB}{136,34,85}
\definecolor{myyellow}{RGB}{153,153,51}
\definecolor{mylightyellow}{RGB}{221,204,119}

\usepackage{microtype}

\usepackage[backend=biber,mcite,subentry,citestyle=numeric-comp,bibstyle=numericbringhurst,autopunct=false,sorting=none,sortcites=false,natbib=false,maxnames=8,minnames=8,giveninits=true,block=space,hyperref=true,defernumbers=false,useprefix=true,language=british]{biblatex}
\renewcommand*{\finalnamedelim}{, }
\setcounter{biburlnumpenalty}{1}
\setcounter{biburlucpenalty}{0}
\setcounter{biburllcpenalty}{1}
\DeclareDelimFormat{multicitedelim}{\addsemicolon\space}
\DeclareDelimFormat{postnotedelim}{\space}
\addbibresource{portamanabib.bib}
\renewcommand{\bibfont}{\footnotesize}
%\defbibheading{bibliography}[\bibname]{\section*{#1}\addcontentsline{toc}{section}{#1}%\markboth{#1}{#1}
%}
\newcommand*{\citep}{\parencites}
\newcommand*{\citey}{\parencites*}
\renewcommand*{\cite}{\citep}
\providecommand{\href}[2]{#2}
\providecommand{\eprint}[2]{\texttt{\href{#1}{#2}}}
\newcommand*{\amp}{\&}

\def\arxivp{}
\def\mparcp{}
\def\philscip{}
\def\biorxivp{}
\newcommand*{\arxivsi}{\texttt{arXiv} eprints available at \url{http://arxiv.org/}.\\}
\newcommand*{\mparcsi}{\texttt{mp\_arc} eprints available at \url{http://www.ma.utexas.edu/mp_arc/}.\\}
\newcommand*{\philscisi}{\texttt{philsci} eprints available at \url{http://philsci-archive.pitt.edu/}.\\}
\newcommand*{\biorxivsi}{\texttt{bioRxiv} eprints available at \url{http://biorxiv.org/}.\\}
\newcommand*{\arxiveprint}[1]{\global\def\arxivp{\arxivsi}%\citeauthor{0arxivcite}\addtocategory{ifarchcit}{0arxivcite}%eprint
\texttt{\urlalt{http://arxiv.org/abs/#1}{arXiv:\hspace{0pt}#1}}%
%\texttt{\href{http://arxiv.org/abs/#1}{\protect\url{arXiv:#1}}}%
%\renewcommand{\arxivnote}{\texttt{arXiv} eprints available at \url{http://arxiv.org/}.}
}
\newcommand*{\mparceprint}[1]{\global\def\mparcp{\mparcsi}%\citeauthor{0mparccite}\addtocategory{ifarchcit}{0mparccite}%eprint
\texttt{\urlalt{http://www.ma.utexas.edu/mp_arc-bin/mpa?yn=#1}{mp\_arc:\hspace{0pt}#1}}%
%\texttt{\href{http://www.ma.utexas.edu/mp_arc-bin/mpa?yn=#1}{\protect\url{mp_arc:#1}}}%
%\providecommand{\mparcnote}{\texttt{mp_arc} eprints available at \url{http://www.ma.utexas.edu/mp_arc/}.}
}
\newcommand*{\philscieprint}[1]{\global\def\philscip{\philscisi}%\citeauthor{0philscicite}\addtocategory{ifarchcit}{0philscicite}%eprint
\texttt{\urlalt{http://philsci-archive.pitt.edu/archive/#1}{PhilSci:\hspace{0pt}#1}}%
%\texttt{\href{http://philsci-archive.pitt.edu/archive/#1}{\protect\url{PhilSci:#1}}}%
%\providecommand{\mparcnote}{\texttt{philsci} eprints available at \url{http://philsci-archive.pitt.edu/}.}
}
\newcommand*{\biorxiveprint}[1]{\global\def\biorxivp{\biorxivsi}%\citeauthor{0arxivcite}\addtocategory{ifarchcit}{0arxivcite}%eprint
\texttt{\urlalt{http://biorxiv.org/content/early/#1}{bioRxiv:\hspace{0pt}#1}}%
%\texttt{\href{http://arxiv.org/abs/#1}{\protect\url{arXiv:#1}}}%
%\renewcommand{\arxivnote}{\texttt{arXiv} eprints available at \url{http://arxiv.org/}.}
}

\usepackage{graphicx}

\usepackage{hyperref}
\usepackage[depth=4]{bookmark}
\hypersetup{colorlinks=true,bookmarksnumbered,pdfborder={0 0 0.25},citebordercolor={0.2 0.1333 0.5333},%bluish
citecolor=myblue,linkbordercolor={0.0667 0.4667 0.2},%greenish
linkcolor=myred,urlbordercolor={0.5333 0.1333 0.3333},%reddish
urlcolor=mygreen,breaklinks=true,pdftitle={\pdftitle},pdfauthor={\pdfauthor}}

% \usepackage[vertfit=local]{breakurl}% only for arXiv
\providecommand*{\urlalt}{\href}

\usepackage{wrapfig}
\usepackage{datetime2}
\DTMnewdatestyle{mydate}%
{% definitions
\renewcommand*{\DTMdisplaydate}[4]{%
\number##3\ \DTMenglishmonthname{##2} ##1}%
\renewcommand*{\DTMDisplaydate}{\DTMdisplaydate}%
}
\DTMsetdatestyle{mydate}

% handling orphan/widow lines:
\clubpenalty=10000
\widowpenalty=10000
\raggedbottom

\selectlanguage{british}\frenchspacing

% to compile a camera-ready version, add the [final] option, e.g.:
% \usepackage[final]{nips_2017}

%\usepackage[utf8]{inputenc} % allow utf-8 input
%\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
%\usepackage{hyperref}       % hyperlinks
%\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
%\usepackage{amsfonts}       % blackboard math symbols
%\usepackage{nicefrac}       % compact symbols for 1/2, etc.
%\usepackage{microtype}      % microtypography

% @@@@@@@@@@ new macros @@@@@@@@@@
% Common ones - uncomment as needed
%\providecommand{\nequiv}{\not\equiv}
%\providecommand{\coloneqq}{\mathrel{\mathop:}=}
%\providecommand{\eqqcolon}{=\mathrel{\mathop:}}
%\providecommand{\varprod}{\prod}
%\newcommand*{\de}{\partialup}%partial diff
%\newcommand*{\pu}{\piup}%constant pi
\newcommand*{\delt}{\deltaup}%Kronecker, Dirac
%\newcommand*{\eps}{\varepsilonup}%Levi-Civita, Heaviside
%\newcommand*{\riem}{\zetaup}%Riemann zeta
%\providecommand{\degree}{\textdegree}% degree
%\newcommand*{\celsius}{\textcelsius}% degree Celsius
%\newcommand*{\micro}{\textmu}% degree Celsius
%\newcommand*{\I}{\mathrm{i}}%imaginary unit
%\newcommand*{\e}{\mathrm{e}}%Neper
%\newcommand*{\di}{\mathrm{d}}%differential
%\newcommand*{\Di}{\mathrm{D}}%capital differential
%\newcommand*{\planckc}{\hslash}
%\newcommand*{\avogn}{N_{\textrm{A}}}
%\newcommand*{\NN}{\bm{\mathrm{N}}}
%\newcommand*{\ZZ}{\bm{\mathrm{Z}}}
%\newcommand*{\QQ}{\bm{\mathrm{Q}}}
\newcommand*{\RR}{\bm{\mathrm{R}}}
%\newcommand*{\CC}{\bm{\mathrm{C}}}
%\newcommand*{\nabl}{\bm{\nabla}}%nabla
%\DeclareMathOperator{\lb}{lb}%base 2 log
%\DeclareMathOperator{\tr}{tr}%trace
%\DeclareMathOperator{\card}{card}%cardinality
%\DeclareMathOperator{\im}{Im}%im part
%\DeclareMathOperator{\re}{Re}%re part
%\DeclareMathOperator{\sgn}{sgn}%signum
%\DeclareMathOperator{\ent}{ent}%integer less or equal to
%\DeclareMathOperator{\Ord}{O}%same order as
%\DeclareMathOperator{\ord}{o}%lower order than
%\newcommand*{\incr}{\triangle}%finite increment
\newcommand*{\defd}{\coloneqq}
\newcommand*{\defs}{\eqqcolon}
%\newcommand*{\Land}{\bigwedge}
%\newcommand*{\Lor}{\bigvee}
%\newcommand*{\lland}{\mathbin{\ \land\ }}
%\newcommand*{\llor}{\mathbin{\ \lor\ }}
%\newcommand*{\lonlyif}{\mathbin{\Rightarrow}}%implies
\newcommand*{\limplies}{\mathbin{\Rightarrow}}%implies
%\newcommand*{\mimplies}{\Rightarrow}%implies
%\newcommand*{\liff}{\mathbin{\Leftrightarrow}}%if and only if
\renewcommand*{\|}{\mathpunct{|}}%conditional sign (in probabilities)
%\newcommand*{\lcond}{\mathpunct{|\ }}%conditional sign (in probabilities)
%\newcommand*{\bigcond}{\mathpunct{\big|}}%conditional sign (in probabilities)
%\newcommand*{\lbigcond}{\mathpunct{\big|\ }}%conditional sign (in probabilities)
%\newcommand*{\suchthat}{\mid}%{\mathpunct{|}}%such that (eg in sets)
%\newcommand*{\bigst}{\mathpunct{\big|}}%such that (eg in sets)
%\newcommand*{\with}{\colon}%with (list of indices)
%\newcommand*{\mul}{\times}%multiplication
%\newcommand*{\inn}{\cdot}%inner product
\newcommand*{\dotv}{\mathord{\,\cdot\,}}%variable place
%\newcommand*{\comp}{\circ}%composition of functions
%\newcommand*{\con}{\mathbin{:}}%scal prod of tensors
%\newcommand*{\equi}{\sim}%equivalent to 
%\newcommand*{\corr}{\mathrel{\hat{=}}}%corresponds to
%\providecommand{\varparallel}{\ensuremath{\mathbin{/\mkern-7mu/}}}%parallel (tentative symbol)
\renewcommand{\le}{\leqslant}%less or equal
\renewcommand{\ge}{\geqslant}%greater or equal
\DeclarePairedDelimiter\clcl{[}{]}
%\DeclarePairedDelimiter\clop{[}{[}
%\DeclarePairedDelimiter\opcl{]}{]}
%\DeclarePairedDelimiter\opop{]}{[}
%\DeclarePairedDelimiter\abs{\lvert}{\rvert}
%\DeclarePairedDelimiter\norm{\lVert}{\rVert}
\DeclarePairedDelimiter\set{\{}{\}}
%\DeclareMathOperator{\pr}{P}%probability
\newcommand*{\pf}{\mathrm{p}}%probability
\newcommand*{\p}{\mathrm{P}}%probability
%\newcommand*{\tf}{\mathrm{T}}%probability
%\renewcommand*{\|}{\|}
%\newcommand*{\+}{\lor}
%\renewcommand{\*}{\land}
\newcommand*{\sect}{\S}% Sect.~
\newcommand*{\sects}{\S\S}% Sect.~
\newcommand*{\chap}{ch.}%
\newcommand*{\chaps}{chs}%
%\newcommand*{\fn}{fn}%
\newcommand*{\eqn}{eq.}%
\newcommand*{\eqns}{eqs}%
\newcommand*{\fig}{fig.}%
\newcommand*{\figs}{figs}%
%\newcommand*{\vs}{{vs.}}
%\newcommand*{\etc}{{etc.}}
\newcommand*{\ie}{{i.e.}}
%\newcommand*{\ca}{{c.}}
%\newcommand*{\Ie}{{I.e.}}
%\newcommand*{\Eg}{{E.g.}}
\newcommand*{\eg}{{e.g.}}
%\newcommand*{\viz}{{viz}}
\newcommand*{\cf}{{cf.}}
%\newcommand*{\Cf}{{Cf.}}
%\newcommand*{\vd}{{v.}}
%\newcommand*{\Vd}{{V.}}
\newcommand*{\etal}{{et al.}}
%\newcommand*{\etsim}{{et sim.}}
%\newcommand*{\ibid}{{ibid.}}
%\newcommand*{\sic}{{sic}}
%\newcommand*{\id}{\mathte{I}}%id matrix
%\newcommand*{\nbd}{\nobreakdash}%
%\newcommand*{\bd}{\hspace{0pt}}%
%\def\hy{-\penalty0\hskip0pt\relax}
\newcommand*{\labelbis}[1]{\tag*{(\ref{#1})$_\text{r}$}}
%\newcommand*{\mathbox}[2][.8]{\parbox[t]{#1\columnwidth}{#2}}
%\newcommand*{\zerob}[1]{\makebox[0pt][l]{#1}}
%\newcommand*{\tprod}{\mathop{\textstyle\prod}\nolimits}
%\newcommand*{\tsum}{\mathop{\textstyle\sum}\nolimits}
%\newcommand*{\tint}{\begingroup\textstyle\int\endgroup\nolimits}
%\newcommand*{\tland}{\mathop{\textstyle\bigwedge}\nolimits}
%\newcommand*{\tlor}{\mathop{\textstyle\bigvee}\nolimits}
%\newcommand*{\sprod}{\mathop{\textstyle\prod}}
%\newcommand*{\ssum}{\mathop{\textstyle\sum}}
%\newcommand*{\sint}{\begingroup\textstyle\int\endgroup}
%\newcommand*{\sland}{\mathop{\textstyle\bigwedge}}
%\newcommand*{\slor}{\mathop{\textstyle\bigvee}}
\newcommand*{\T}{^\intercal}%transpose
\newcommand*{\E}{\mathrm{E}}
\DeclarePairedDelimiter\expp{(}{)}
\newcommand*{\expe}{\E\expp}%round
\newcommand*{\expeb}{\E\clcl}%square
%%\newcommand*{\QEM}%{\textnormal{$\Box$}}%{\ding{167}}
%\newcommand*{\qem}{\leavevmode\unskip\penalty9999 \hbox{}\nobreak\hfill
%\quad\hbox{\QEM}}

\newtheoremstyle{simple}%
{}%.5\baselineskip plusminus.2\baselineskip % \parsep
{}%
{\footnotesize}%
{}%
{}%
{}%
{0pt}%
{}%
\theoremstyle{simple}
\newtheorem*{simplenote}{}

\definecolor{notecolour}{RGB}{68,170,153}
\newcommand*{\puzzle}{{\fontencoding{U}\fontfamily{fontawesometwo}\selectfont\symbol{225}}}
\newcommand{\mynote}[1]{ {\color{notecolour}\puzzle\ #1}}
\newcommand*{\widebar}[1]{{\mkern1.5mu\skew{2}\overline{\mkern-1.5mu#1\mkern-1.5mu}\mkern 1.5mu}}

\newcommand*{\av}{\widebar} %pop average
\newcommand*{\sav}{\widebar} %subpop average

\newcommand*{\ypp}{\varPi}

\newcommand*{\yvX}{\varSigma}
\newcommand*{\yvR}{\sigma}
\newcommand*{\yvx}{S}
\newcommand*{\yvr}{s}
\newcommand*{\yvN}{\nu}

\newcommand*{\yx}{\bm{\yvx}}%subpop state
\newcommand*{\yxs}{\sav{\yx}}%subpop av state
\newcommand*{\yX}{\bm{\yvX}}%pop state
\newcommand*{\yXf}{\av{\yX}}%pop av state
\newcommand*{\yxxs}{\sav{\yx\yx}}%subpop av state
\newcommand*{\yXXf}{\av{\yX\yX}}%subpop av state
\newcommand*{\yr}{\bm{\yvr}}%subpop value
\newcommand*{\yrs}{\sav{\yr}}%subpop av value
\newcommand*{\yR}{\bm{\yvR}}%pop value
\newcommand*{\yRf}{\av{\yR}}%pop av value
%conditional assumptions
\newcommand*{\yH}{\varIota}
\newcommand*{\yD}{\varDelta}
\newcommand*{\yHa}{\varIota_\textrm{p}}
\newcommand*{\yHb}{\varIota_\textrm{s}}
\newcommand*{\yHc}{\varKappa}
\newcommand*{\yHd}{\varUpsilon}
\newcommand*{\yHp}{\varIota'}
\newcommand*{\yHi}{\varIota''}
%Lagr multipliers
\newcommand*{\yf}{\bm{f}}
\newcommand*{\yc}{\bm{c}}
\newcommand*{\yL}{\bm{\varLambda}}
\newcommand*{\yl}{\bm{\lambda}}
\newcommand*{\yk}{\zeta}
\newcommand*{\yK}{\varZeta}
%entropies
\newcommand*{\ysh}{H}
\newcommand*{\ybu}{H_{\text{B}}}

\newcommand*{\prop}[1]{`#1'}

\newcommand*{\mee}{maximum entropy}
\newcommand*{\me}{maximum-entropy}

%@@@@@@@@@@ new macros end @@@@@@@@@@

\selectlanguage{british}\frenchspacing

\title{\pdftitle}

% The \author macro works with any number of authors. There are two
% commands used to separate the names and addresses of multiple
% authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to
% break the lines. Using \AND forces a line break at that point. So,
% if LaTeX puts 3 of 4 authors names on the first line, and the last
% on the second line, try using \AND instead of \And before the third
% author name.

\author{
  P.G.L. Porta Mana\\\texttt{piero.mana@ntnu.org}
  \And
  V. Rostami\\\texttt{v.rostami@fz-juelich.de}
  \AND
  E. Torre\\\texttt{torre@ibk.baug.ethz.ch}
    \And
Y. Roudi\\\texttt{yasser.roudi@ntnu.no}
  % David S.~Hippocampus\thanks{Use footnote for providing further
  %   information about author (webpage, alternative
  %   address)---\emph{not} for acknowledging funding agencies.} \\
  % Department of Computer Science\\
  % Cranberry-Lemon University\\
  % Pittsburgh, PA 15213 \\
  % \texttt{hippo@cs.cranberry-lemon.edu} \\
  %% examples of more authors
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \AND
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
}

\begin{document}
% \nipsfinalcopy is no longer used

\maketitle

\begin{abstract}
  \mynote{to be rewritten} This note shows that the \me\ method can be
  applied to a representative sample from a neuronal population along two
  different routes: (1) apply to the sample; or (2) apply to the population
  and marginalize to the sample. These two routes give inequivalent
  results. Which route should be chosen? Some arguments are presented in
  favour of the second. The note also touches upon probability formulae of
  representative sampling and discusses their possible meanings, a
  discussion that may be useful for sampling problems in neuroscience.
  % The abstract must be limited to one paragraph.
\end{abstract}

%\section{Implicit assumptions in the maximum-entropy method}
\section{Introduction: \me\ and sampling in neuroscience}

Suppose that we have recorded the firing activity of a hundred neurons,
sampled from a particular brain area. What are we to do with such data?
Gerstein, Perkel, \amp\ Dayhoff \cite{gersteinetal1985} posed this question
very tersely (our emphasis):
\begin{quote}
  The principal conceptual problems are (1) \emph{defining cooperativity or
    functional grouping} among neurons and (2) \emph{formulating
    quantitative criteria} for recognizing and characterizing such
  cooperativity.
\end{quote}
These questions have a long history, of course; see \eg\ the 1966 review by
Moore \etal\ \cite{mooreetal1966}. % This scenario is a concrete possibility
% thanks to recent electrophysiological techniques \cite{berenyietal2014}.
The neuroscientific literature has offered several mathematical definitions
of \enquote{cooperativity} or \enquote{functional grouping} and criteria to
quantify it.

One such quantitative criterion relies on the \me\ or relative-\me\ method
\cite{jaynes1957,jaynes1963,hobsonetal1973,sivia1996_r2006,meadetal1984}.
This criterion has been used in neuroscience at least since the 1990s,
applied to data recorded from brain areas as diverse as retina and motor
cortex
\cite{mackay1991,martignonetal1995,bohteetal2000,amarietal2003,schneidmanetal2006,shlensetal2006,mackeetal2009b,roudietal2009c,tkaciketal2009,gerwinnetal2009,mackeetal2011,mackeetal2011b,ganmoretal2011,granotatedgietal2013,tkaciketal2014b,moraetal2015,shimazakietal2015},
and it has been subjected to mathematical and conceptual scrutiny
\cite{tkaciketal2006,roudietal2009,roudietal2009b,barreiroetal2010b,mackeetal2013,rostamietal2016_r2017}.

\enquote{Cooperativity} can quantified and characterized with \me\ method
in several ways. The simplest way roughly proceeds along the following
steps. Consider the recorded activity of a sample of $n$ neurons.
\begin{enumerate}
\item The activity of each neuron, a continuous signal, is binned in time
  and binarized in intensity, and thus transformed into a sequence of
  digits \enquote{$0$}s (inactive) and \enquote{$1$}s
  \cite[\cf][]{caianiello1961,caianiello1986}.

  Let the variable $\yvx_i(t)\in\set{0,1}$ denote the activity of the $i$th
  sampled neuron at time bin $t$. The average activity at that bin is
  therefore $\yxs(t) \defd \sum_i \yvx_i(t)/n$. If we count the number of
  distinct pairs of active neurons at that bin we combinatorially find
  $\tbinom{n\yxs(t)}{2}\equiv n\yxs(t)\,[n\yxs(t)-1]/2$. There can be at
  most $\tbinom{n}{2}$ simultaneously active pairs, so the average pair
  activity is
  $\av{\yx \yx}(t) \defd \tbinom{n}{2}^{-1}\tbinom{n\yxs(t)}{2}$. With some
  combinatorics we see that the average activity of $m$-tuples of neurons
  is
  \begin{equation}
    \label{eq:products_intermsof_average}
    \av{\underbrace{\yx\dotsm\yx}_{\text{$m$ terms}}}
    = \binom{n}{m}^{-1}\binom{n\yxs(t)}{m}.
  \end{equation}
  For brevity let us agree to simply call \enquote{activity} the average
  $\yxs$, \enquote{pair-activity} the average $\av{\yx \yx}$, and so on.

\item Construct a sequence of relative-\me\ distributions for the activity
  $\yxs$, using this sequence of constraints:
  \begin{itemize}
  \item the time average of the activities $\yxs(t)$;
  \item the time averages of the activities $\yxs(t)$ and pair-activities
    $\av{\yx \yx}(t)$;
  \item \ldots and so on, including the time averages of $\yxs(t)$ up to
    $\av{\yx\dotsm\yx}(t)$ with $k$ terms.
  \end{itemize}
  Call the resulting distributions $p_1(\yxs), p_2(\yxs),\dotsc,p_k(\yxs)$.
  The time-bin dependence is now absent because these distributions can be
  interpreted as referring to any one of the time bins $t$, or to a new
  time bin containing new data.

  We also have the empirical frequency distribution of the total activity,
  $f(\yxs)$, counted from the time bins.

\item Now compare the distributions above with one another and with the
  frequency distribution, using some probability-space distance like the
  relative entropy or discrimination information
  \citep{kullback1987,jaynes1963,hobson1969,hobsonetal1973}. If we find,
  say, that such distance is very high between $p_1$ and $f$, very low
  between $p_2$ and $f$, and is more or less the same between all $p_m$ and
  $f$ for $m \ge 2$, then we can say that there is a \enquote{pairwise
    cooperativity}, and that any higher-order cooperativity is just a
  reflection or consequence of the pairwise one. The reason is that the
  information from higher-order simultaneous activities did not lead to
  appreciable changes in the distribution obtained from pair activities.
\end{enumerate}
The protocol above needs to be made precise by specifying various
parameters, such as the width of the time bins or the probability
distance used.

We hurry to say that the description just given is just \emph{one} way to
quantify and characterize cooperativity and functional grouping, not
\emph{the only} way. It can surely be criticized from many points of view.
Yet, it is quantitative and bears a more precise meaning than an undefined,
vague notion of \enquote{cooperativity}. Two persons who apply this
procedure to the same data will obtain the same numbers. Different
protocols can be based on the \me\ method, for instance protocols that take
into account the activities or pair activities of specific neurons rather
than population averages, or even protocols that take into account time
dependence.

The purpose of the present work is not to assess the merits of \me\ methods
with respect to other methods. Its main purpose is to show that there is a
problem in the way the \me\ method itself, as sketched above, is applied to
the activity of the recorded neurons. We believe that this problem is at
the root of some quirks about this method that were pointed out in the
literature \cite{roudietal2009b}. This problems extends also to more
complex versions of the method, possibly except version that use
\enquote{hidden} neurons. The problem stems from the fact that the recorded
neurons are a \emph{sample} from a larger, unrecorded population.

Suppose that the neurons were recorded with electrodes covering an area of
some square millimetres \cite[\cf][]{berenyietal2014}. This recording is a
sample of the activity of the neuronal population under the recording
device; such population can amount to tens of thousands of neurons
\cite{abeles1991}. We could even consider the recorded neurons as a sample
of a brain area more extended than the recording device.

The characterization of the cooperativity of the recorded sample would have
little meaning if we did not expect its results to generalize to a larger,
unrecorded population -- at the very least that under the recording device.
In other words, we expect that the conclusions drawn with the \me\ methods
about the sampled neurons should somehow extrapolate to unrecorded neurons
in some larger area, from which the recorded neurons were sampled. In
statistical terms we are assuming that the recorded neurons are a
\emph{representative sample} of some larger neuronal population.
Probability theory tells us how to make inferences from a sample to the
larger population from which it is sampled \parentext{see references
  below}.

Here is the problem. We can apply the \me\ method to the sample, as
described in the above protocol, to generate probability distributions for
the activity of the sample. But, if our sample is representative of a
larger population, we can apply the \me\ method to the larger (unrecorded)
population, using the constraints obtained from the recorded sample, and
generate a probability distribution for the larger population. The
distribution for the sample is then obtained by marginalization. Either
application seems to have some commendable features. However, \emph{the
  distributions obtained from these two applications differ}. Which choice
is most meaningful?

In this paper we develop the second way of applying the \me, at the level
of the larger population, and show that its results differ from the
application at the sample level. We also consider the case where the size
of the larger population is unknown.

To apply the \me\ method to the larger, unsampled population, it is
necessary to use probability relations relevant to sampling
\cites{ghoshetal1997}[parts~I, VI]{freedmanetal1978_r2007}[\chap~8]{gelmanetal1995_r2014}[\chap~3]{jaynes1994_r2003}.
The relations we present are well-known in survey sampling and in the
pedagogic problem of drawing from an urn without replacement, yet they are
somewhat hard to find explicitly written in the neuroscientific literature,
so they may be of interest on their own.

\iffalse In the final discussion we present some arguments in favour of one choice
in our dilemma, but we do not mean that to be our final answer. This note,
and its dilemma,  also have the more general purpose of reminding our
readers the importance of asking \emph{what is the question we're trying to
  answer with probability theory?} and \emph{which assumptions do we want
  or need to make?}

Our mathematical analysis pertains to neurons modelled as binary units at
one specific instant or short window in time. It is possible to similarly
discuss multi-state neuron models and population dynamics; but for
simplicity neurons are here assumed to be in a fixed, active \enquote{$1$}
or inactive \enquote{$0$} state; and evolution, change, time correlations,
and similar concepts do not concern us. We consider \me\ models that
constrain various kinds of sample or population averages; these models are
often called \enquote{homogeneous}. The final discussion touches upon
\enquote{inhomogeneous} models as well.
\fi

The notation and terminology in this note follows ISO and ANSI standards
\citep{iso1993,ieee1993,nist1995,iso2006,iso2006b} but for the use of the
comma \enquote{,} to denote logical conjunction. Probability notation
follows Jaynes \citep{jaynes1994_r2003}. By \enquote{probability} we mean a
degree of belief which \enquote{would be agreed by all rational men if
  there were any rational men} \cite{good1966}.

\section{Notation: population, sample, probabilities}
\label{sec:setup}

We have a population of $\yvN$ binary neurons. We assume that they can be
distinguished, by their spike shapes for example; but other details, like
their locations, are unknown. The neurons have a joint state
$(\yvX_1,\dotsc,\yvX_\yvN) \defs \yX$ having fixed but unknown binary values
$(\yvR_1,\dotsc,\yvR_\yvN)\defs \yR \in \set{0,1}^\yvN$; the \enquote{$\yvX$} denote the
variables, and \enquote{$\yvR$} their values. A particular sample of $n$ neurons
from this population has joint state $(\yvx_1, \dotsc, \yvx_n) \defs \yx$ having
fixed binary values $(\yvr_1, \dotsc, \yvr_n)\defs\yr \in \set{0,1}^n$. We will
consider various averages of the population and the sample. For this
purpose we introduce a population-averaging operator $\sav{\dotv}$ defined
by
\begin{equation}
  \begin{gathered}
    \av{\yX} \defd \tfrac{1}{\yvN} (\yvX_1 + \yvX_2 + \dotsb + \yvX_\yvN),
    % & \sav{\yx} &\defd \tfrac{1}{n} (x_1 + x_2 + \dotsb + x_n),
   \qquad% \\[2\jot]
    \av{\yX \yX} \defd \tbinom{\yvN}{2}^{-1} (\yvX_1 \yvX_2 + \yvX_1 \yvX_3 + \dotsb +
    \yvX_{\yvN-1} \yvX_\yvN),
    % & \sav{\yx \yx} &\defd\tbinom{n}{2}^{-1} (x_1 x_2 + x_1 x_3 + \dotsb
    % + x_{n-1} x_n),
    \\[2\jot]
    \av{\yX\yX\yX} \defd \tbinom{\yvN}{3}^{-1} (\yvX_1 \yvX_2 \yvX_3 +% \yvX_1 \yvX_2 \yvX_4
    \dotsb + \yvX_{\yvN-2} \yvX_{\yvN-1}\yvX_\yvN),
    \qquad\text{and so on.}
    % & \sav{\yx\yx\yx} &\defd \tbinom{n}{3}^{-1} (x_1 x_2 x_3
    % + %x_1 x_2 x_4
    % \dotsb + x_{n-2} x_{n-1} x_n),
  \end{gathered}
\end{equation}
These formulae say that $\yXf$ is the fraction of active neurons, $\yXXf$
the fraction of simultaneously active pairs out of all $\binom{\yvN}{2}$
pairs, $\av{\yX\yX\yX}$ the fraction of simultaneously active triplets, and
so on. Products of states like $\yvX_i \dotsm \yvX_j$ also have values in
$\set{0,1}$; from this we can combinatorially  prove that
\begin{equation}
  \label{eq:products_intermsof_average}
  \av{\underbrace{\yX\dotsm\yX}_{\text{$m$ factors}}}
  = \binom{\yvN}{m}^{-1}\binom{\yvN\yXf}{m}.
\end{equation}
Analogous formulae  hold for quantities like $\yx$, $\yR$, $\yr$.

\iffalse
\begin{simplenote}
  \mynote{Remove this} A sample can be \enquote{random} in two different
  ways. Consider the population of neurons named $\set{1,2,3}$ for example.
  The sample $\set{1,3}$ is random because neurons $1$ and $3$ were chosen
  according to an unknown process. The sample
  $\set{\textrm{X}_1, \textrm{X}_2}$ is random because we do not know the
  identity of neurons $\textrm{X}_1$ and $\textrm{X}_2$. Given the
  probability for the population, the probabilities for the states of these
  two samples are different: the former is obtained from marginalization, the
  latter by symmetrization and marginalization.
\end{simplenote}
\fi

Our uncertainty about the actual state of the population is completely
expressed by the joint probability distribution
\begin{equation}
  \label{eq:joint_plaus}
  \p(\yvX_1=\yvR_1, \yvX_2=\yvR_2, \dotsc, \yvX_\yvN=\yvR_\yvN \| \yHc) \quad\text{or}\quad
\p(\yX =\yR \| \yHc), \quad \yR \in \set{0,1}^\yvN,
\end{equation}
where $\yHc$ denotes our state of knowledge, \ie\ the evidence and
assumptions backing this particular probability assignment. Our uncertainty
about the state of the sample is likewise expressed by
\begin{equation}
  \label{eq:sample_plaus}
  \p(\yvx_1=\yvr_1, \yvx_2=\yvr_2, \dotsc, \yvx_n=\yvr_n \| \yHc) \quad\text{or}\quad
\p(\yx =\yr \| \yHc), \quad \yr \in \set{0,1}^n.
\end{equation}

\section{Probability: relations between population and  sample}
\label{sec:prob_samples}

The theory of statistical sampling is covered in many excellent texts, for
example Ghosh \amp\ Meeden \citey{ghoshetal1997} or Freedman, Pisani, \amp\
Purves \citey[parts~I, VI]{freedmanetal1978_r2007}; summaries can be found
in Gelman \etal\ \citey[\chap~8]{gelmanetal1995_r2014} and Jaynes
\citey[\chap~3]{jaynes1994_r2003}.

We need to make an initial probability assignment for the state of the
population before any experimental observations are
made. %, no matter what kinds of predictions we are interested in.
This initial assignment will be modified by our experimental observations,
which can involve only a sample of the population. Our initial probability
assignment should reflect that samples are somehow representative of the
whole population. We consider here a state of knowledge that express
this representativeness.

In this state of knowledge, denoted $\yH$, we know that the neurons in the
population are biologically or functionally similar, for example in
morphology or the kind of input or output they receive or give. But we are
completely ignorant about the physical details of the individual neurons.
Our ignorance is therefore symmetric under permutations of neuron
identities. This ignorance is represented by a probability distribution
that is symmetric under permutations of neuron identities; such a
distribution is usually called \emph{finitely exchangeable}
\cites{ericson1969}[\chap~1]{ghoshetal1997}. We stress that this
probability assignment is just an expression of the symmetry of our
\emph{ignorance} about the state of the population, not an expression of
some biologic or physical identity of the neurons.

The \emph{representation theorem for finite exchangeability} states that
the symmetric distribution in the state of knowledge $\yH$ must obey
\begin{equation}
  \label{eq:joint_plaus_N_homog}
  \p(\yX = \yR \|  \yH) \equiv
  \p(\yX = \yR \| \yXf = \yRf, \yH)\,
  \p(\yXf=\yRf \| \yH) =
  \binom{\yvN}{\yvN\yRf}^{-1} \p(\yXf=\yRf \| \yH),
\end{equation}
the latter being the probability for the population average $\yX$. A sum is
only apparently missing in the central term: its summands $\yXf=A$ are all
zero except for $A=\yRf$. Proof of this theorem and generalizations to
non-binary and continuum cases are given by de~Finetti
\cite{definetti1959b}, Kendall \cite{kendall1967}, Ericson
\cite{ericson1976}, Diaconis \amp\ Freedman
\cite{diaconis1977,diaconisetal1980}, Heath \amp\ Sudderth
\cite{heathetal1976}. This theorem is intuitive: owing to symmetry, we must
assign equal probabilities to all states with $\yvN\yRf$ active neurons.

By marginalization we obtain the probability for the state of the sample:
\begin{gather}
  \label{eq:marginal}
  \p(\yx = \yr \| \yH) = \binom{n}{n\yrs}^{-1} \p(\yxs=\yrs \| \yH),
  \\
  \shortintertext{with}
  \label{eq:subpop_average}
  \p(\yxs=\yrs \| \yH) = \sum_{\yvN \yRf =0}^{\yvN}
    \p(\yxs = \yrs \|\yXf=\yRf, \yH)\,
  \p(\yXf=\yRf \| \yH),
  \\[2\jot]
  \label{eq:conditional_hypergeometric}
  \p(\yxs = \yrs \|\yXf=\yRf, \yH)=
  \binom{n}{n\yrs}\binom{\yvN-n}{\yvN \yRf-n\yrs}\binom{\yvN}{\yvN \yRf}^{-1}
  \defs \ypp(\yrs\|\yRf).
\end{gather}
The conditional probability in the last formula is a hypergeometric
distribution $\ypp(\yrs\|\yRf)$, typical of \enquote{drawing without
  replacement} problems. The combinatorial proof of the formulae above is
in fact the same as for this class of problems
\cites[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.
Our initial symmetric ignorance should intuitively also apply to the
sample; the probability for the state of the
sample~\eqref{eq:subpop_average} indeed satisfies the representation
theorem~\eqref{eq:joint_plaus_N_homog} as well.

The conditional probability
$\p(\yxs = \yrs \|\yXf=\yRf, \yH) \equiv \ypp(\yrs,\yRf)$ relates the
spaces of the sample average $\yXf\in\set{0,\dotsc,\yvN}$ and of the
population average $\yxs\in\set{0,\dotsc,n}$. It is a coarsening projector
of any probability for $\yXf$ onto a marginal probability for $\yxs$, as
\eqn~\eqref{eq:subpop_average} shows. For each function $f$ of the sample
average $\yxs$, the  hypergeometric distribution 
$\ypp(\yrs,\yRf)$ also allows us to construct a function $f^*$ of the
population average $\yXf$
\begin{equation}
  \label{eq:pullback_P}
    f^*(\yXf) \defd \sum_{n\yrs=0}^n f(\yrs)\ypp(\yrs\|\yXf),
%  \ypp^*\colon f(\yxs=\dotv) \to f^*(\yXf=\dotv) \defd \sum_{N\yrs=0}^n
%  f(\yrs)\ypp(\yrs\|\dotv),
\end{equation}
that has the same expectation as $f$:
\begin{equation}
  \expeb{f^*(\yXf)}
  \equiv
  \sum_{\yvN\yRf=0}^\yvN f^*(\yRf)\,\p(\yXf=\yRf \| \yH)
%  =
%  \sum_{N\yRf=0}^N \sum_{n\yrs=0}^n f(\yrs)\ypp(\yrs\|\yRf)\,\p(\yXf=\yRf \| \yH)
  =
  \sum_{n\yrs=0}^n f(\yrs)\,\p(\yxs=\yrs \| \yH)
  \equiv  \expeb{f(\yxs)}.
\end{equation}
The middle equality uses the combination of formulae~\eqref{eq:pullback_P}
and~\eqref{eq:subpop_average}.

\begin{figure}[!t]
\centering
\includegraphics[width=0.9\linewidth]{pop_sample_projection3.pdf}%
\caption{Log-plot of the hypergeometric distribution
  $%\p(\yxs = \yrs \|\yXf=\yRf, \yH)=
\ypp(\yrs\|\yRf) \defd  \raisebox{0pt}[0pt][1ex]{}\binom{n}{n\yrs}\binom{\yvN-n}{\yvN \yRf-n\yrs}\binom{\yvN}{\yvN \yRf}^{-1}$ for $\yvN=5000$, $n=200$. (Band artifacts may appear in the
  colourbar depending on your \textsc{pdf} viewer.)}
\label{fig:hypergeom_proj}
\end{figure}%hypergeometric_identity.nb
A look at a plot of the hypergeometric distribution $\ypp(\yrs\|\yRf)$, see
\fig~\ref{fig:hypergeom_proj}, reveals that it is a sort of \enquote{fuzzy
  identity matrix} between the $\yXf$-space $\set{0,\dotsc,\yvN}$ and
$\yxs$-space $\set{0,\dotsc,n}$. When $n=\yvN$ it is the identity matrix. We
thus have that
\begin{equation}
  \label{eq:roughly_equal_nN}
  \p(\yxs=a) \approx \p(\yXf=a),\qquad
\expeb{f(\yxs)} \approx \expeb{f(\yXf)}.
\end{equation}
These are only very approximate equalities: they may miss important
features of the two probability distributions. In the next section we will
in fact emphasize their differences. If the distribution for the population
average $\yXf$ is bimodal, for example, the bimodality can be lost in the
distribution for the sample average $\yxs$, owing to the coarsening effect
of $\ypp(\yrs\|\yRf)$.

The approximate equalities above, on the other hand, express the fact that
\emph{our uncertainty about the sample is representative of our uncertainty
  about the population and about other samples}, and vice versa. These
approximate equalities are not generally valid for other probability
distributions for the population. They should be contrasted with the limits
$\p(\yxs=a) \to \p(\yXf=a)$ and $\expeb{f(\yxs)} \to \expeb{f(\yXf)}$, as
$n\to \yvN$, do: these limits are trivially, universally valid because the
sample becomes the full population as $n\to \yvN$. In particular, these limits
hold even in cases where the conditional probability
$\p(\yxs = \yrs \|\yXf=\yRf)$ is not a fuzzy identity, our uncertainties
about sample and about population can differ wildly, and the approximate
equalities~\eqref{eq:roughly_equal_nN} do not hold.

\bigskip

% The projection~\eqref{eq:projection_P} and pull-back~\eqref{eq:pullback_P}
% have the property
% \begin{equation}
%   \label{eq:expectation_pullback}
%   \expeb{f(\yxs) \|\yH} = \expeb{f^*(\yXf)\|\yH}
%   \qquad\text{or}\qquad
%   \sum_{n\yrs=0}^n f(\yrs)\,p_*(\yxs=\yrs) = \sum_{\yvN\yRf=0}^\yvN f^*(\yRf)\,p(\yXf=\yRf).
% \end{equation}
For averaged products, $\sav{\underbrace{\yx\dots\yx}_{\text{$m$
      factors}}}\equiv\binom{n\yxs}{m}/\binom{n}{m}$,
formulae~\eqref{eq:pullback_P} yield the useful equality
\begin{equation}
  \label{eq:expe_products}
%  \label{eq:pullback_m_expectations}
%  \ypp^*\colon \sav{\underbrace{\yx \dotsm \yx}_{\text{$m$ factors}}}
%  \mapsto
    % \bigl( \sav{\underbrace{\yx \dotsm \yx}_{\text{$m$ factors}}}\bigr)^* =
    % \av{\underbrace{\yX \dotsm \yX}_{\text{$m$ factors}}},
    % \\[2\jot]
    % \expe{\sav{\underbrace{\yx \dotsm \yx}_{\text{$m$ factors}}} \| \yH} =
    % \expe{\av{\underbrace{\yX \dotsm \yX}_{\text{$m$ factors}}} \| \yH}
  \expe{\sav{\underbrace{\yx \dotsm \yx}_{\text{$m$ factors}}} \| \yH}
\equiv
    % \\[2\jot]
    % {}
    \binom{n}{m}^{-1}
    % \frac{(n-m)!}{n!}
    \sum_{\mathclap{n \yrs=0}}^n %m!\,
    \binom{n \yrs}{m}\, \p(\yxs=\yrs \| \yH)
    % \sum_{N \yRf=0}^N %m!\,
    % \binom{N-m}{N \yRf-m} \binom{N}{N \yRf}^{-1} \p(\yXf=\yRf \| \yH)
    = \binom{\yvN}{m}^{-1}
    % \frac{(N-m)!}{N!}
    \sum_{\mathclap{\yvN \yRf=0}}^\yvN %m!\,
    \binom{\yvN \yRf}{m}\, \p(\yXf=\yRf \| \yH)
\equiv    \expe{\av{\underbrace{\yX \dotsm \yX}_{\text{$m$ factors}}} \| \yH}.
\end{equation}
The proof of the middle equality uses the expression for the $m$th
factorial moment of the hypergeometric distribution \cite{potts1953}. Thus,
in the state of knowledge $\yH$ the expectations of the averages of
activity products \emph{are initially the same for the sample and for the
  full population}. Similar relations can be found for the raw moments
$\expe{\yxs^m}$ and $\expe{\yXf^m}$, which can be written in terms of the
product expectations via \eqn~\eqref{eq:products_intermsof_average}.

\section{Maximum-entropy at the population level with constraints at the
  sample level}
\label{sec:specific_initial_probability}

The probability formulae
\eqref{eq:joint_plaus_N_homog}--\eqref{eq:conditional_hypergeometric} are
constraints on our initial probability assignment, but do not determine it
numerically. The probability $\p(\yXf=\yRf \| \yH)$ for the population
average needs to be numerically specified, and by marginalization
\eqref{eq:subpop_average} it will determine that of the sample average,
$\p(\yxs=\yrs \| \yH)$. If we numerically specify the latter, the former is
not completely specified, because \eqn~\eqref{eq:subpop_average} linearly
constrains $\yvN+1$ unknowns by only $n+1$ equations.

We may want to specify the probability by enforcing the sample expectations
of several functions to have specific values, for example
$\expe{\yxs}=c_1$, $\expe{\yxxs}=c_2$. This is still an underdetermined
problem: several distributions can have the same desired expectations, as
clear from \eqns~\eqref{eq:expe_products}.

The \me\ method is brought into play to solve this indeterminacy. It
selects one distribution, purported to be \enquote{maximally noncommittal},
among those that have the desired expectations. But here's a dilemma: the
expectation formulae~\eqref{eq:pullback_P} allow us to apply the method to
find the probability of the population $\p(\yXf=\yRf \| \yH)$, or of the
sample $\p(\yxs=\yrs \| \yH)$. \emph{The two applications, however, are
  inequivalent}. They lead to numerically different distributions for the
sample average $\p(\yxs=\yrs \| \yH)$.

Suppose we want to constrain the sample expectations of a vector function
$\yf = (f_1,\dotsc,f_m)$ to the vector values $\yc=(c_1,\dotsc,c_m)$, that
is, $\expeb{\yf(\yxs)}=\yc$. Application of \me\
\cite{sivia1996_r2006,meadetal1984} at the population level, denoted by
$\yHa$, gives
\begin{equation}
  \label{eq:app_maxent_pop}
  \p(\yXf = \yRf \| \yHa)  = \yK\,
  \binom{\yvN}{\yvN \yRf}\,\exp\Bigl[\yL\T
  \sum_{n\yrs=0}^n \yf(\yrs)\ypp(\yrs\|\yRf)\Bigr],
\end{equation}
and then by marginalization~\eqref{eq:marginal}
\begin{equation}
  \label{eq:app_maxent_pop_marg}
  \p(\yxs = \yrs \| \yHa)  = \yK\, 
  \sum_{\yvN\yRf=0}^\yvN \ypp(\yrs \|\yRf)
\,
  \binom{\yvN}{\yvN \yRf}\,\exp\Bigl[\yL\T
  \sum_{n\yrs=0}^n \yf(\yrs)\ypp(\yrs\|\yRf)\Bigr],
\end{equation}
where $\yK$ is a normalization constant and
$\yL\T=(\varLambda_1,\dotsc,\varLambda_m)\T$ are Lagrange multipliers such
that
\begin{equation}
  \label{eq:app_maxent_pop_constraints}
\yc = \yK\sum_{n\yrs=0}^n \sum_{\yvN\yRf=0}^\yvN \yf(\yrs) \ypp(\yrs \|\yRf)
\,
  \binom{\yvN}{\yvN \yRf}\,\exp\Bigl[\yL\T
  \sum_{n\yrs=0}^n \yf(\yrs)\ypp(\yrs\|\yRf)\Bigr].
\end{equation}
Application of \me\ at the sample level, denoted by $\yHb$, gives
\begin{equation}
  \label{eq:app_maxent_sample}
  \p(\yxs=\yrs \|\yHb)
  =\yk\,
  \binom{n}{n\yrs}
  \exp[\yl\T \yf(\yrs)]
\end{equation}
where $\yk$ is a normalization constant and $\yl\T$ are Lagrange
multipliers such that
\begin{equation}
  \label{eq:app_maxent_sample_constraints}
\yc = \yk\sum_{n\yrs=0}^n  \yf(\yrs) \binom{n}{n\yrs}
  \exp[\yl\T \yf(\yrs)].
\end{equation}

The probabilities for the sample average obtained from application at the
population level~\eqref{eq:app_maxent_pop_marg} and at the sample level
\eqref{eq:app_maxent_sample} should be approximately equal, by our previous
observation about representativity~\eqref{eq:roughly_equal_nN} and also by
the fact that they must satisfy the same expectations for $\yf$.

Yet they cannot be exactly equal, because their equality would require the
Lagrange multipliers $\yL$ and $\yl$ to satisfy the constraint equations
\eqref{eq:app_maxent_pop_constraints},
\eqref{eq:app_maxent_sample_constraints}, and also
$\p(\yxs = \yrs \| \yHa) = \p(\yxs = \yrs \| \yHb)$; that is, $2m+n$
equations (one normalization is taken care of) in $2m$ unknowns. A solution can
exist, if at all, only for very special choices of constraints functions
$\yf$ and values $\yc$.

The sample distribution obtained from \me\ at the sample level will
therefore likely miss important features present in the one obtained at the
population level, like additional modes or particular tail behaviour.
\begin{figure}[!t]
\centering
\includegraphics[width=0.9\linewidth]{different_maxent_pop_sample_200_realdata_2mom.pdf}%
\caption{Linear and log-plots of $\p(\yxs = \yrs)$ constructed by \me\ at the population
  level followed by sample marginalization (blue triangles),
  \eqn~\eqref{eq:app_maxent_pop_marg}, and at the sample level (red
  circles), \eqn~\eqref{eq:app_maxent_sample}, with $\yvN=5000$,
  $n=200$, constraints $\expe{\yxs}=0.0478$, $\expe{\yxxs}=0.00257$.}
\label{fig:diff_maxent_pop_sample}
\end{figure}%maxent_pop_or_sample.nb
% \end{wrapfigure}
\begin{figure}[!t]
\centering
\includegraphics[width=0.9\linewidth]{different_maxent_pop_sample_200_realdata_4mom.pdf}
% 
\caption{Linear and log-plots of $\p(\yxs = \yrs)$ constructed by \me\ at
  the population level followed by sample marginalization (blue triangles),
  \eqn~\eqref{eq:app_maxent_pop_marg}, and at the sample level (red
  circles), \eqn~\eqref{eq:app_maxent_sample}, with $\yvN=5000$,
  $n=200$, constraints $\expe{\yxs}=0.0478$, $\expe{\yxxs}=0.00257$,
  $\expe{\sav{\yx\yx\yx}}=1.48\times10^{-4}$,
  $\expe{\sav{\yx\yx\yx\yx}}=8.81\times 10^{-6}$.}
\label{fig:diff_maxent_pop_sample_realdata}
\end{figure}%maxent_pop_or_sample.nb
We show two examples of this discrepancy in
\figs~\ref{fig:diff_maxent_pop_sample}
and~\ref{fig:diff_maxent_pop_sample_realdata}, for $\yvN=5000$, $n=200$, and
constraint functions of the form
$\yf(\yxs)=(\yxs,\yxxs,\dotsc)\equiv\bigl( \yxs,
\raisebox{0pt}[0pt][1ex]{}\binom{n\yxs}{2}/\binom{n}{2},\dotsc \bigr)$,
equivalent to moments constraints. The constraint values used in these
examples, reported in the figure captions, have neurobiologically realistic
values \cite{rostamietal2016_r2017}.

In the first example the constraint functions are $\expe{\yxs}$ and
$\expe{\yxxs}$. The distribution obtained at the sample level is broader
than the one obtained at the population level; the tails of the two
distributions are very different.

The second
example uses two additional constraint functions $\expe{\sav{\yx\yx\yx}}$,
$\expe{\sav{\yx\yx\yx\yx}}$. The distribution obtained at the population
level has two modes, replaced by only one in the distribution obtained at
the sample level; the tails are very different also in this case.

\bigskip

How should we apply the \me\ method then? on the sample or on the
population? Which application is \enquote{maximally noncommittal}?

\section{Discussion}
\label{sec:discussion}

The question that closed the preceding section cannot receive a categorical
answer. An optimal answer can only be given case by case, depending on the
computational power available, on which inferences we are trying to make,
on which assumptions we need or want to make, and those we wish to avoid.

\iffalse We want to stress, in fact, the
extreme importance and priority of the following two questions:
\begin{itemize}
\item Of which quantities do we need to quantify the uncertainty? and why?
\item What are the assumptions we need or want to make to answer the
  question above?
\end{itemize}\fi

\iffalse The first purpose of this note is indeed to remind ourselves that
probability models require careful scrutiny, because they can rest on
hidden assumptions that we do not want to make or that contradict others we
are making, and we may not be aware of this. The dilemma generated by the
\me\ method and the assumption of representative sampling is an example.
\fi

The tricky point is this. The \me\ application at the population level and
the application at the sample level give different results; they are two
different statistical models. The former model clearly assumes, by
construction, the existence of a larger population from which the sample is
taken. What does the latter model assume in this respect? is it
\enquote{unassuming}, as often claimed in the literature? or does it
actually assume that \emph{no} larger population exists? In the latter case
it would not be correct to use this model in our problem.

A perfunctory intuitive reasoning seems insufficient for clarifying this
point. Let's express it in the language of the probability calculus.
Suppose we do not know whether the sample is really part of a larger
population: we do not know whether $\yvN=n$ or how large $\yvN$ is otherwise.
Call this state of ignorance $\yHd$. In the probability calculus this
ignorance about $\yvN$ is expressed by assigning a probability distribution
$\p(\yvN \| \yHd)$ that vanishes if $\yvN<n$, since we know that $\yvN\ge n$; see
Good \cite{good1965,good1967b} and Rissanen \cite{rissanen1983} for
examples of such distributions over the integers. Maintaining our
assumption of symmetric ignorance, probability assignments that do not
assume a specific value of $\yvN$ are then obtained via multiplication of all
$\yvN$-dependent probabilities by $\p(\yvN \| \yHd)$ and subsequent
marginalization over $\yvN$. Technically speaking, $\yvN$ becomes a
\emph{nuisance parameter}
\cite{jaynes1994_r2003,lindley1965b_r2008,bernardoetal1994_r2000}. The
probability obtained from \me\ at the population level,
\eqn~\eqref{eq:app_maxent_pop_marg}, then generalizes to
\begin{equation}
  \label{eq:app_maxent_pop_marg_unknown_N}
  \p(\yxs = \yrs \| \yHd)  = \sum_\yvN \biggl\{\yK_\yvN\, 
  \sum_{\yvN\yRf=0}^\yvN \ypp_\yvN(\yrs \|\yRf)
\,
  \binom{\yvN}{\yvN \yRf}\,\exp\Bigl[{\yL_\yvN}\T
  \sum_{n\yrs=0}^n \yf(\yrs)\ypp_\yvN(\yrs\|\yRf)\Bigr]\biggr\}
  \,\p(\yvN \| \yHd),
\end{equation}
where $\yvN$-dependencies have been made explicit. This is a formidable
expression. But our question, \enquote{is the usual \me\ at the sample
  level~\eqref{eq:app_maxent_sample} unassuming with regard to the
  existence of a larger population?}, translates now into the precise
mathematical question: \enquote{are the
distributions~\eqref{eq:app_maxent_pop_marg_unknown_N} and
\eqref{eq:app_maxent_sample} equal for some choice of $\p(\yvN \| \yHd)$, with
$\p(\yvN \| \yHd)\ne 0$ for $\yvN>n$?}. We leave this mathematical problem for
future work. Note, however, that this equality is satisfied if
$\p(\yvN=1 \| \yHd)=1$, which means that the usual \me\ model can also be
interpreted as assuming that \emph{no} larger population exists.

\iffalse
The former application must therefore implicitly assume that no
larger population exist, or that the neurons of the sample are isolated
from their environment.

The \me\ application at the population level gives in fact, by
construction, a distribution with lower Shannon relative entropy than the
one given by the application at the sample level: both satisfy the same
constraints, but only the latter is obtained via direct maximization of the
entropy. The lower entropy of the distribution obtained from application at
the population level is clearly visible from its higher sharpness in
\fig~\ref{fig:diff_maxent_pop_sample}. Such distribution therefore
excludes, probabilistically speaking, more sample states $\yx$ than the
distribution obtained at the sample level does. A lower entropy signals the
presence of additional assumptions or constraints. In this case the
assumption is that the sample is part of a larger population, and our
ignorance about the sample should be representative of our ignorance about
a larger population. This assumption affects our predictions.
\fi

We find the \me\ model constructed at the population level very natural and
preferable. After all, physical models of neuronal networks usually include
some sort of external input to the neurons as well, mimicking their
embedding in a larger network. The sample distribution given by the \me\
model at the population level, when used as a reference distribution for
surprise analysis, may reveal features in a dataset that were unnoticed by
the standard \me\ model. The question remains of how to specify $\yvN$,
though. We have tacitly intended $\yvN$ as the size of the largest
biologically or functionally homogeneous population from which our sample
was recorded. It could be the amount of neurons in a functional brain area,
for example the primary visual cortex, for which $\yvN\sim 10^8$
\cite{leubaetal1994}. For large $\yvN$ -- unfortunately we are not yet able to
translate this \enquote{large} into a numeric order of magnitude -- the
final distribution becomes independent of $\yvN$, and continuous
approximations become available.

\iffalse
The quandary we found in the present note has at least two bright sides.
The first is that the two different \me\ applications may give numerically
similar results in some cases. One could thus be used as an approximation
of the other, depending on the circumstances -- \eg\ if probability tails
are unimportant. See \fig~\ref{fig:diff_maxent_pop_sample_realdata} for an
example.
\begin{figure}[!t]
\centering
\includegraphics[width=0.95\columnwidth]{different_maxent_pop_sample_200_realdata_both.pdf}
% 
\caption{Example of discrepancy in $\p(\yxs = \yrs \| \yH)$ constructed by
  \me\ at the population level followed by sample marginalization (blue
  triangles), and at the sample level (red circles). The constraints are
  the same in either case: $\expe{\yxs \|\yH}=0.045$,
  $\expe{\yxxs \|\yH}=0.0025$, and $\yvN=5000$, $n=200$.}
\label{fig:diff_maxent_pop_sample_realdata}
\end{figure}%maxent_pop_or_sample.nb
The second is that \fi

The possibility of using two different distributions is not a physical
contradiction. Similar situations arise in statistical mechanics. It is
known that if a system is described by a \me\ Gibbs state, its subsystems
need not be \cite{maesetal1999}. A dilemma quite similar to ours also
appears in the statistical description of the final state of a
non-equilibrium process starting and ending in two equilibrium states: we
can describe our knowledge about the final state either by a Gibbs
distribution, or by the distribution obtained from the Liouville evolution
of the Gibbs distribution assigned to the initial state. The two
descriptions differ -- even though the final \emph{physical} state is
obviously exactly the same \cite[\sect~4]{jaynes1985d_r1993}. The two
descriptions differs because in one case we can make sharper predictions
about the state thanks to our knowledge of its preceding dynamics. In this
example, though, both distributions are usually immensely sharp and
practically lead to the same predictions. In the neuroscientific
applications considered in this note the difference in predictions may be
relevant instead.

Our analysis touched only constraints of the sample average,
$\expeb{f(\yxs)}$. The corresponding models are usually called
\enquote{homogeneous} in the literature. Purely \enquote{inhomogeneous}
models have also been used
\cite{schneidmanetal2006,shlensetal2006,roudietal2009b}, in which
expectations for individual neurons or groups of neurons are constrained,
for example $\expe{\yvx_2}$ or $\expe{\yvx_1x_8x_9}$. A short computation shows
that the \me\ method with this kind of constraints gives the same result
whether applied at the sample or at the population level: the states of any
unconstrained neurons marginalize out. This is understandable: expressing
different uncertainties about, say, neurons $2$ and $5$ we are breaking the
symmetry of our uncertainty, which thus cannot be representative of other
neurons in the sample or in the population. Inhomogeneous models, however,
require enormous computational power for large sample sizes; homogeneous
models therefore retain their importance. Our analysis and dilemma also
persist for hybrid homogeneous-inhomogeneous models
\cite{tkaciketal2014b,shimazakietal2015}.

%\mynote{Importance of the formulae relating sample and population.}

% \mynote{Discussion of case when individual $\yvX_i\dotsm \yvX_j$ are constrained.
% Likely breakdown of \me\ in this case \cite{portamana2009}. Hints at full
% Bayesian treatment of which \me\ is a limit? \cite[***]{mackay1995_r2003}}

% do not change font sizes (except perhaps in the \textbf{References}
% section; see below).

\subsubsection*{Acknowledgments}

To be added after review.
\iffalse PGLPM thanks Mari \amp\ Miri for continuous encouragement and affection,
Buster Keaton for filling life with awe and inspiration, and the developers
and maintainers of \LaTeX, Emacs, AUC\TeX, MiK\TeX, arXiv, biorXiv,
PhilSci, Hal archives, Python, Inkscape, Sci-Hub for making a free and
unfiltered scientific exchange possible.\fi

% Use unnumbered third level headings for the acknowledgments. All
% acknowledgments go at the end of the paper. Do not include
% acknowledgments in the anonymized submission, only in the final paper.

%\section*{References}
\renewcommand*{\bibname}{References}
\defbibheading{bibliography}[\bibname]{\section*{#1}\addcontentsline{toc}{section}{#1}%\markboth{#1}{#1}
}

% References follow the acknowledgments. Use unnumbered first-level
% heading for the references. Any choice of citation style is acceptable
% as long as you are consistent. It is permissible to reduce the font
% size to \verb+small+ (9 point) when listing the references. {\bf
%   Remember that you can go over 8 pages as long as the subsequent ones contain
%   \emph{only} cited references.}
% \medskip

% \small

% [1] Alexander, J.A.\ \& Mozer, M.C.\ (1995) Template-based algorithms
% for connectionist rule extraction. In G.\ Tesauro, D.S.\ Touretzky and
% T.K.\ Leen (eds.), {\it Advances in Neural Information Processing
%   Systems 7}, pp.\ 609--616. Cambridge, MA: MIT Press.

% [2] Bower, J.M.\ \& Beeman, D.\ (1995) {\it The Book of GENESIS:
%   Exploring Realistic Neural Models with the GEneral NEural SImulation
%   System.}  New York: TELOS/Springer--Verlag.

% [3] Hasselmo, M.E., Schnell, E.\ \& Barkai, E.\ (1995) Dynamics of
% learning and recall at excitatory recurrent synapses and cholinergic
% modulation in rat hippocampal region CA3. {\it Journal of
%   Neuroscience} {\bf 15}(7):5249-5262.
\defbibnote{postnote}{\small\par\medskip\noindent{\footnotesize% Note:
\arxivp \mparcp \philscip \biorxivp}%
}

\newcommand*{\citein}[2][]{\textnormal{\textcite[#1]{#2}}%\addtocategory{extras}{#2}
}
\newcommand*{\citebi}[2][]{ref.\ \citep[#1]{#2}%\addtocategory{extras}{#2}
}
\newcommand*{\subtitleproc}[1]{}

\printbibliography[postnote=postnote]

\iffalse
\newrefsection
\bigskip
{\centering\mynote{*** From here down old text to be discarded ***}\par}
\section{***}

\begin{align}
  \expe{\yvx_{i_1} \dotsm \yvx_{i_m} \| \yH}
&=
\expe{\sav{\underbrace{\yx \dotsm \yx}_{\text{$m$ factors}}} \| \yH}
=
\expe{\av{\underbrace{\yX \dotsm \yX}_{\text{$m$ factors}}} \| \yH},
\\
\intertext{and has an explicit expression in terms of $Q$:}
  \begin{split}
\expe{\yvx_{i_1} \dotsm \yvx_{i_m} \| \yH}
&=
\binom{\yvN}{m}^{-1}
%\frac{(N-m)!}{N!}
\sum_{\yvN A=0}^\yvN %m!\, 
\binom{\yvN A}{m}\, Q(A)
\\
&\equiv
%\frac{(N-m)!}{N!}
\sum_{\yvN A=0}^\yvN %m!\, 
\binom{\yvN-m}{\yvN A-m}
\binom{\yvN}{\yvN A}^{-1} Q(A).
\end{split}
\end{align}
\end{subequations}

What do we mean by saying that we can learn something about the population
by observing the sample? At the very least we mean that our uncertainty
about the rest of the population can change upon observing the sample. That
is, the probability for the state $\yy$ of the rest of the population is
conditionally dependent on the sample's values:
\begin{equation}
  \label{eq:conditionally_dependent_sample}
  \p(\yy =\ys \| \yx=\yr,\yH) \ne \p(\yy =\ys \|\yH)
  \qquad\text{for some $\ys,\yr$},
\end{equation}
This also means that the probability for the population cannot be
factorized into that of the sample times that of the complement.

What do we mean when we say that the sample is representative of the
population? It means that we expect some collective properties of the
sample, like the fraction of active neurons, or the fraction of
simultaneously active pairs, to be roughly equal to those of the full
population.

***

For example, $\yX$ can
represent the state of a population at a particular time. We call the
neurons \enquote{units} to lend some generality to our discussion. We shall
make statements about the whole population of $\yvN$ units and about a
subpopulation of $n$ units; the word \enquote{population} will always refer
to the \emph{whole} population. The subpopulation states and their values
are denoted by lowercase letters: $(\yvx_1, \dotsc, \yvx_n)\equiv\yx$ and
$(\yvr_1, \dotsc, \yvr_n)\equiv\yr$; but note that $\yvx_i \equiv \yvX_{j_i}$ and
$\yvr_i \equiv \yvR_{j_i}$ for some distinct $j_1,\dotsc,j_n$. We shall also make
statements about the population-averaged state, or \emph{population
  average}:
\begin{align}
  \label{noeq:population-av}
   \yXf &\defd (\yvX_1 + \dotsb + \yvX_\yvN)/\yvN,
\\
\intertext{and the subpopulation-averaged state, or \emph{subpopulation average}:}
\label{noeq:subpopulation-av}
\yxs &\defd (\yvx_1 + \dotsc + \yvx_n)/n.
\end{align}
The quantities $\yvN\yXf$ and $n\yxs$ represent the total number of active
units in the population and the subpopulation. Quantities like $\yRf$ and $\yrs$
are defined analogously. The averaging operators $\sav{\dotv}$ and
$\av{\dotv}$ are also extended to averages of $\tbinom{n}{m}$ or
$\tbinom{\yvN}{m}$ products of $m$ states; \eg,
\begin{align}
\av{\yX \yX} &\defd
\tbinom{\yvN}{2}^{-1} (\yvX_1 \yvX_2 + \yvX_1 \yvX_3  + \dotsb +  \yvX_{\yvN-1} \yvX_\yvN),
\\[2\jot]
\sav{\yx\yx\yx} &\defd
\tbinom{n}{3}^{-1} (\yvx_1 \yvx_2 \yvx_3 + \yvx_1 \yvx_2 \yvx_4 + \dotsb + \yvx_{n-2} \yvx_{n-1} \yvx_n),
\end{align}
and so on.

\subsection{Assumptions}
\label{nosec:assumptions}

Our uncertainty about the population state is represented by the joint
probability distribution of the individual states, from which we can
derive all other probabilities of interest. We denote it by
\begin{equation}
  \label{noeq:joint_plaus}
  \p(\yvX_1=\yvR_1, \yvX_2=\yvR_2, \dotsc, \yvX_\yvN=\yvR_\yvN \| \yH) \quad\text{or}\quad
\p(\yX =\yR \| \yH).
\end{equation}
Such probability is conditional on our state of knowledge, \ie\ the
evidence and assumptions backing our probability assignments, denoted by
the proposition $\yH$.

In the present discussion, $\yH$ is a state of knowledge that leads to two
specific properties in our probability assignments:

\medskip
\begin{enumerate}%[wide,label=$\yH$\arabic*.]
\item \emph{Permutation symmetry}, expressed as the invariance of the
  joint distribution~\eqref{noeq:joint_plaus} under arbitrary permutations of
  the units's labels:
\begin{multline}
  \label{noeq:homog_math}
  \p(\yvX_1=\yvR_1, \yvX_2=\yvR_2, \dotsc, \yvX_\yvN=\yvR_\yvN \| \yH) ={}
\\ 
\shoveright{\p(\yvX_1=\yvR_{\pi(1)}, \yvX_2=\yvR_{\pi(2)}, \dotsc, \yvX_\yvN=\yvR_{\pi(\yvN)} \| \yH)}
\\
\text{for any permutation $\pi$}.
\end{multline}
This property can reflect two very different states of knowledge: physical
homogeneity of the population, or symmetry in our ignorance about the population.
This property is called \emph{finite exchangeability} in the Bayesian
literature and its basis, consequences, and alternatives to it are discussed
in \sect~***%\ref{nosec:stoch_symm}.

\medskip

\item The population average $\yXf$ has a particular distribution $Q$:
\begin{equation}
  \label{noeq:pop_average_Q}
  \p(\yXf = A \| \yH)
% \equiv \pr(x_1+ \dotsb + x_N = N A \| \varEta) 
=  Q(A),
\qquad
A \in \set*{0,\tfrac{1}{\yvN},\tfrac{2}{\yvN},\dotsc,1}.
\end{equation}
For the moment we are not concerned about the specific form of $Q$ and
about how it was assigned: it could, \eg, arise from maximum-entropy
arguments
\citep[\eg:][]{jaynes1957,jaynes1963,good1963,jaynes1967,aczeletal1975,jaynes1979b,vancampenhoutetal1981,sivia1990,fangetal1997,bretthorst2013}
used with data on the population.
\end{enumerate}

\subsection{Formulae}
%\section{\ding{41} {\itshape The mathematical relations} \ding{41}}
%\section*{3.\hspace{\stretch{1}}The mathematical relations\hspace{\stretch{1}}\mbox{}}
%\addcontentsline{toc}{section}{The mathematical relations}
\label{nosec:main_formulae}

The state of knowledge $\yH$ has the following six (not independent) main
consequences for our probability assignments:

\begin{enumerate}%[wide,label=\textbf{\emph{\Roman*}}.]
  \medskip\item\label{noitem:tot_plaus}Probability for the population state:
\begin{equation}
  \label{noeq:joint_plaus_N_homog}
  \p(\yX = \yR \| H) = \binom{\yvN}{\yvN\yRf}^{-1} Q(\yRf).
\end{equation}

\medskip
\item\label{noitem:marginal}Probability for the state $\yx$ of any subpopulation
  of $n$ units:
\begin{equation}
  \label{noeq:marginal}
    \p(\yx = \yr \| \yH) =
\sum_{\yvN A =0}^{\yvN} 
\binom{\yvN-n}{\yvN A-n\yrs}\binom{\yvN}{\yvN A}^{-1}\,
Q(A).
\end{equation}
Note that the only summands contributing to this sum are those for which
$n\yrs \le NA \le \yvN$; the others are zero because by definition
$\tbinom{M}{y}=0$ if $y<0$. This remark applies to all the sums of this
kind in the rest of this Note.

\medskip
\item\label{noitem:conditional}Probability for the subpopulation
  state conditional on a population state:
\begin{equation}
  \label{noeq:conditional_total_state}
  \p(\yx = \yr \| \yX = \yR, \yH)
=
\binom{\yvN-n}{\yvN \yRf-n \yrs}.
\end{equation}

\medskip
\item\label{noitem:marginal_average}Probability for the
  subpopulation average $\yxs$:
\begin{multline}
  \label{noeq:subpop_average}
  \p(\yxs = a  \| \yH) = \binom{n}{n a}
 \sum_{\yvN A=0}^{\yvN}
\binom{\yvN-n}{\yvN A-n a}\binom{\yvN}{\yvN A}^{-1}\,
Q(A),
\\
a \in \set*{0,\tfrac{1}{n},\tfrac{2}{n},\dotsc,1}.
\end{multline}

\medskip
\item\label{noitem:conditional_average}Probability for the subpopulation
  average conditional on the population average:
\begin{equation}
  \label{noeq:conditional_prob}
  \p(\yxs = a \| \yXf = A, \yH)
=
\binom{n}{n a}\binom{\yvN-n}{\yvN A-n a}\binom{\yvN}{\yvN A}^{-1}.
\end{equation}

\medskip
\item\label{noitem:moments}The product of the states of any $m$ distinct
  units from a given subpopulation,
  \begin{equation*}
    \yvx_{i_1} \yvx_{i_2} \dotsm \yvx_{i_m},
    \qquad 1\le i_1 < i_2 < \dotsb < i_m \le n
  \end{equation*}
  has an expectation equal to that of the subpopulation average of such
  products, is independent of the subpopulation size $n$:
\begin{subequations}
\label{noeq:expe_products}
\begin{align}
  \expe{\yvx_{i_1} \dotsm \yvx_{i_m} \| \yH}
&=
\expe{\sav{\underbrace{\yx \dotsm \yx}_{\text{$m$ factors}}} \| \yH}
=
\expe{\av{\underbrace{\yX \dotsm \yX}_{\text{$m$ factors}}} \| \yH},
\\
\intertext{and has an explicit expression in terms of $Q$:}
  \begin{split}
\expe{\yvx_{i_1} \dotsm \yvx_{i_m} \| \yH}
&=
\binom{\yvN}{m}^{-1}
%\frac{(N-m)!}{N!}
\sum_{\yvN A=0}^\yvN %m!\, 
\binom{\yvN A}{m}\, Q(A)
\\
&\equiv
%\frac{(N-m)!}{N!}
\sum_{\yvN A=0}^\yvN %m!\, 
\binom{\yvN-m}{\yvN A-m}
\binom{\yvN}{\yvN A}^{-1} Q(A).
\end{split}
\end{align}
\end{subequations}
\end{enumerate}

\medskip

A useful relation connects the expectation of a
product~\eqref{noeq:expe_products} and the $m$th \emph{factorial moment}
\citep{potts1953} of the probability distributions for the averages. The
$m$th factorial moment of the subpopulation average $\yxs$ is defined by
\begin{equation}
  \label{noeq:fact_mom_defin}
\vphantom{\underbrace{\yxs}_{m}}\expeb*{
\smash{\underbrace{n\yxs\,(n\yxs-1)\dotsm \bigl(n\yxs-(m-1)\bigr)}_{\text{$m$ factors}}}
\| \yH}
\equiv \expeb*{\frac{(n\yxs)!}{(n\yxs-m)!} \| \yH},
\end{equation}
an analogous definition holding for $\yXf$. We have that
\begin{equation}
  \label{noeq:eq_factmom_prod}
\expe{\yvx_{i_1} \dotsm \yvx_{i_m} \| \yH}
=
  \tfrac{(n-m)!}{n!}\expeb*{\frac{(n\yxs)!}{(n\yxs-m)!} \| \yH}
=
  \tfrac{(\yvN-m)!}{\yvN!}\expeb*{\frac{(\yvN\yXf)!}{(\yvN\yXf-m)!} \| \yH}.
\end{equation}

As a consequence of the above relation, the first three moments of the
probability distributions $\p(\yxs=a\|\yH)$ and $\p(\yXf=A\|\yH)$,
are related by
\begin{subequations}\label{noeq:moments}
  \begin{align}
    \expe{\yxs \| \yH} &= \expe{\yXf \| \yH},
    \\[2\jot]
    \begin{split}
    \expe{\yxs^2 \| \yH} &= \expe{\yXf \| \yH}
                                  \frac{\yvN-n}{(\yvN-1)n}
 +{}
\\&\quad\quad \expe{\yXf^2 \| \yH}\frac{\yvN\,(n-1)}{(\yvN-1)n},
\end{split}
    \\[2\jot]
    \begin{split}
    \expe{\yxs^3 \| \yH} &=
                              \expe{\yXf \| \yH}\frac{(\yvN-n)(\yvN-2n)}{(\yvN-1)(\yvN-2)n^2}+{}
                                    \\
                                    &\quad\quad\expe{\yXf^2 \| \yH}
                                    \frac{3N\,(\yvN-n)(n-1)}{(\yvN-1)(\yvN-2)n^2} +{}
                                    \\
                                    &\quad\quad\quad\expe{\yXf^3 \| \yH}
                                    \frac{\yvN^2\,(n-1)(n-2)}{(\yvN-1)(\yvN-2)n^2}.
                                  \end{split}
                                    % \\[2\jot]
                                    % \expe{\yxs^4 \| \varEta} &=
                                    % \!\begin{aligned}[t] &\expe{\yxf
                                    %   \| \varEta}
                                    %   \frac{(N-n)(N^2+N-6Nn+6n^2)}{(N-1)(N-2)(N-3)n^3}+{}
                                    %   \\%[1.5\jot]
                                    %   &\quad\expe{\yxf^2 \| \varEta}
                                    %   \frac{(7N-11n+1)(N-n)(n-1)N}{(N-1)(N-2)(N-3)n^3}+{}
                                    %   \\
                                    %   &\quad\expe{\yxf^3 \| \varEta}
                                    %   \frac{6(n-1)(n-2)(N-n)N^2}{(N-1)(N-2)(N-3)n^3}
                                    %   +{}
                                    %   \\
                                    %   &\quad\expe{\yxf^4 \| \varEta}
                                    %   \frac{(n-1)(n-2)(n-3)N^3}{(N-1)(N-2)(N-3)n^3}.
                                    % \end{aligned}
  \end{align}
\end{subequations}
Relations for higher moments can be obtained recursively from
\eqn~\eqref{noeq:eq_factmom_prod}. In general, this means that the two sets of
first $m$ moments are related by a homogeneous linear transformation,
\begin{equation}
  \label{noeq:lin_transf_moments}
  \expe{\yxs^m \| \yH} = \sum_{l=1}^m M_{ml}(n,\yvN) \expe{\yXf^l \| \yH},
\end{equation}
with a universal, lower-triangular transformation matrix $M_{ml}(n,\yvN)$ that
depends only on $n$, $\yvN$, and the condition of symmetry~\eqref{noeq:homog_math}.

As intuition suggests, we have 
%$\expe{\yxs^m \| \yH} \xrightarrow{n\to \yvN} \expe{\yXf^m \| \yH}$.
\begin{equation}
  \label{noeq:limit_moments}
  \expe{\yxs^m \| \yH} \xrightarrow{n\to \yvN} 
  \expe{\yXf^m \| \yH},
\qquad
  \expe{\yxs^m \| \yH} \xrightarrow{n\to 1} 
  \expe{\yXf \| \yH},
\end{equation}
the latter because ${\yvx_i}^m=\yvx_i$, since states are $\set{0,1}$-valued.

\bigskip

The core of the six mathematical relations above are
\eqns~\eqref{noeq:marginal} and~\eqref{noeq:subpop_average}. The latter
expresses the probability for the subpopulation average as a mixture of
hypergeometric distributions
\cites[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968},
with parameters $\yvN,\yvN\yXf,n$, weighted by the probabilities
$\p(\yXf = A \|\yH)$ \citep[\cf][\sect~4, esp.\ \eqn~(22)]{kendall1967}.
The connection between this mixture representation and the condition of
symmetry~\eqref{noeq:homog_math} is well-known in the Bayesian literature
\citep{kendall1967,definetti1969b,heathetal1976,diaconis1977,diaconisetal1980,jaynes1986c}.

\section{Examples of inferential use of the formulae}

\subsection{From network to subnetwork}
\label{nosec:pop2sub_examples}

Let us illustrate with an example how the probability distribution for the
subnetwork average $\yxs$, determined by \eqn~\eqref{noeq:subpop_average},
changes with the subnetwork size $n$. Choose a network-average distribution
$\p(\yXf=A\| \yH)$ belonging to the exponential family
\cites[\sect~4.5.3]{bernardoetal1994}[see also][]{fortinietal2000}:
\begin{equation}
  \label{noeq:full_pop_av_example_ME}
  \p(\yXf = A \| \yH) = Q(A) \propto
\binom{\yvN}{\yvN A}\exp[
\yl_2\, \yvN A\,(\yvN A-1)/2 + \yl_1\, \yvN A].
\end{equation}
This is the form obtained from the principle of maximum relative entropy
\citep[\eg:][]{jaynes1957,jaynes1963,good1963,jaynes1967,aczeletal1975,jaynes1979b,vancampenhoutetal1981,sivia1990,fangetal1997,bretthorst2013}
with first and second moments as constraints and the reference distribution
$Q_0$ defined by $Q_0(A) = 2^{-\yvN}\binom{\yvN}{\yvN A}$, corresponding to a
uniform probability distribution for the network state $\yX$.
%file: distr_scaling_from_full_population.nb
\begin{figure}[!b]
\centering
\includegraphics[width=0.95\columnwidth]{scaled_subpop_probs.pdf}%
\caption{Probability distributions $\p(\yxs = a \| \yH)$ for different
  subnetwork sizes $n$, obtained from a network probability
  distribution $\p(\yXf = A \| \yH)$ having the maximum-etropy
  form~\eqref{noeq:full_pop_av_example_ME}.}
\label{noscaling_distr}
\end{figure}
\begin{figure}[!b]
\centering
\includegraphics[width=0.95\columnwidth]{scaling_subpop_moments.pdf}%
\caption{Moments of the probability distributions $\p(\yxs =a \|
  \yH)$ as functions of the subnetwork size $n$.}
\label{noscaling_moments}
\end{figure}

The probability distribution of \eqn~\eqref{noeq:full_pop_av_example_ME} is
plotted in \fig~\ref{noscaling_distr}, together with the resulting
subnetwork-average distributions $\p(\yxs = a \|\yH)$, for the case in
which $\yvN = 1000$ units, $\lambda_1=-2.55$, $\lambda_2=0.005$, and
$n=10, 50, 100, 250$. The distributions become broader as $n$ decreases,
and the minimum of the original distribution disappears; at the same time
the finite-difference 
\[\frac{\p(\yxs=a+1/n\|\yH)-\p(\yxs=a\|\yH)}{1/n}\]
presents a sharp jump at this minimum when $n\approx 100$.

To the eye familiar with maximum-entropy distributions, the
subnetwork-average distributions of \fig~\ref{noscaling_distr} do not look
like maximum-entropy ones with second-moment constraints. In fact,
they are not and \emph{cannot} be:
\begin{equation}
  \label{noeq:maxent_form_sub}
  \p(\yxs =a  \| \yH)
\ne\kappa\binom{n}{n a}\exp[\yk_2\, n a\,(n a-1)/2 + \yk_1\, n a]
%\quad\text{(false)}
\end{equation}
for any $\kappa, \yk_1, \yk_2$, unless $n=2$. This impossibility holds more
generally for any number of constraints $m$ and subnetwork size $n$ such
that $m<n$. The reason is simple: suppose we have assigned a
maximum-entropy distribution with $m$ moment constraints as the
distribution for the network average. If we want the same kind of
distribution for a subnetwork of size $n$, we are free to play with $m+1$
parameters (normalization included), but we must also satisfy the $n+1$
equations corresponding to the marginalization~\eqref{noeq:subpop_average}.
This is generally impossible unless $m \ge n$. (Impossibilities of a
similar kind appear in statistical mechanics, see \eg\
ref.~\citep{maesetal1999}.)
%file: test_ME_same_under_scaling.nb

This fact can be significant for recent works
\citep[\eg,][]{schneidmanetal2006,shlensetal2006,tkaciketal2006,marreetal2009,tkaciketal2009,ganmoretal2011,shimazakietal2012,tkaciketal2013,shimazakietal2015}
in which a maximum-entropy probability distribution with second- or
third-moment constraints is assigned to relatively small subnetworks
($n < 200$) of neurons. If we assume that such subnetwork is part of a
larger network, and assume the condition of symmetry~\eqref{noeq:homog_math},
then the larger network \emph{cannot} be assigned a maximum-entropy
distribution with the same number of constraints. Vice versa, if we assign
such a maximum-entropy distribution to the larger network, then none of its
subnetwork of enough large size $n$ can be assigned a similar
maximum-entropy distribution. See ref.~\citep{rostamietal2016_r2017} for a
broader discussion of this fact and of its consequences.

\medskip

The dependence of the first four moments $\expe{\yxs^m \| \yH}$ as a
function of size $n$ is shown in \fig~\ref{noscaling_moments}. The moments
become practically constant when $n \approx 100$ or larger. The
expectations of $m$-tuple products of states
$\expe{\yvx_{i_1}\dotsm \yvx_{i_m} \| \yH}$, proportional to the factorial
moments, are not shown as they do not depend on $n$.

\subsection{From subnetwork to network}
\label{nosec:from_sub_to_full}

We have seen that, given the condition of symmetry~\eqref{noeq:homog_math},
the probability $\p(\yXf = A \|\yH)$ for the network average determines
that of each subnetwork average, $\p(\yxs \|\yH)$, by the
marginalization \eqn~\eqref{noeq:subpop_average}. The reverse is trivially
not true, since \eqn~\eqref{noeq:subpop_average}, as a linear mapping from
$\RR^{\yvN+1}$ to $\RR^{n+1}$, with $\yvN$ larger than $n$, is onto but not into.
Assigning a probability distribution $\p(\yxs = a \|\yH)$ to a
subnetwork average $\yxs$ does not determine a network distribution
$\p(\yXf = A\|\yH)$: it only restricts the set of possible ones; this
set can in principle be determined via linear-programming methods
\citep{hailperin1965,hailperin1984,hailperin1996,hailperin2006,hailperin2011}.
\begin{innote}
  Analogous situations appear in the truth-valued logical calculus: if the
  composite proposition $\varAlpha\limplies \varBeta$ is assigned the
  truth-value \enquote{true}, then assigning $\varAlpha$ the value
  \enquote{true} also determines the value of $\varBeta$, whereas assigning
  $\varBeta$ the value \enquote{true} leaves the value of $\varAlpha$
  undetermined.
\end{innote}
The same linear-programming methods show that any inference from subnetwork
properties to network ones must necessarily start from some assumptions
$\varIota$ that assign a probability distribution
$\p(\yX = \yR \| \varIota)$ for the network states. The approaches to
this task and reformulations of it have become uncountable: they include
exchangeable models, parametric and non-parametric models, hierarchical
models, general linear models, models via sufficiency, maximum-entropy
models, and whatnot
\citep[\eg:][]{jeffreys1931_r1973,jeffreys1939_r2003,jaynes1994_r2003,bernardoetal1994,gelmanetal1995_r2014,ghoshetal1997,kallenberg2005,gregory2005,sivia1996_r2006,ferreiraetal2007,dawid2013,damienetal2013}.
We now show two examples, based on a maximum-entropy approach, that to our
knowledge have not yet been explored in the neuroscientific literature. For
a concrete application see \citep{rostamietal2016b}.

\paragraph{First example: moment constraints for the network.}
\label{nosec:maxent_moments}
Consider a state of knowledge $\yHa$ leading to the following properties:
\begin{enumerate}%[$\yHa$1.]
\item the expectations of the single and pair averages $\yxs$ and
  $\sav{\yx\yx}$ of a particular subnetwork have given values
  \begin{equation}
    \label{noeq:constr_ex1}
    \expe{\yxs \| \yHa} = c_1, \qquad \expe{\sav{\yvx_i \yvx_j}\|\yHa} = c_2;
  \end{equation}
\item the network probability distribution $\p(\yX = \yR \|\yHa)$
  has maximum relative entropy with respect to the uniform one, given the
  constraints above.
\end{enumerate}

\medskip Then the probability distribution for the network conditional on $\yHa$
is completely determined: it satisfies the symmetry
property~\eqref{noeq:homog_math} and is defined by
\begin{multline}
  \label{noeq:pop_distr_maxent1}
  \pf(\yX =\yR \|\yHa) =
\varKappa\exp[\yL_2 \yvN\yRf\,(\yvN\yRf-1)/2 + \yL_1 \yvN\yRf]
\\
\!\begin{aligned}[b]
&\text{with $\varKappa, \yL_m$, such that the distribution is normalized and}
\\
&\varKappa\sum_{\yvN A=0}^\yvN %m!\, 
\binom{\yvN-m}{\yvN A-m}
\exp[\yL_2 \yvN A\,(\yvN A-1)/2 + \yL_1 \yvN A]
= c_m, \quad m=1,2.
\end{aligned}
\end{multline}
We omit the full proof of this statement: it is a standard application of
the maximum-entropy procedure
\citep[\eg:][]{jaynes1957,jaynes1963,good1963,jaynes1967,jaynes1979b,vancampenhoutetal1981,sivia1990,fangetal1997,bretthorst2013},
combined with the equality~\eqref{noeq:expe_products} of subnetwork and
network expectations, \eg
\begin{equation}
  \label{noeq:exploit_eq_expects}
  c_2 = \expe{\sav{\yx\yx}\|\yHa} = 
\binom{\yvN}{2}^{-1}
\sum_{\yvN A=0}^\yvN %m!\, 
\binom{\yvN A}{2} \, \p(\yXf = A\|\yHa),
\end{equation}
and with relations~\eqref{noeq:joint_plaus_N_homog},
\eqref{noeq:conditional_total_state}. This example is easily generalized to
any number $m$ of constraints such that  $m\le n$.

Note again that, as remarked in \sect~\ref{nosec:pop2sub_examples}, the
subnetwork from which the averages in the
expectations~\eqref{noeq:constr_ex1} are calculated has a probability
distribution $\p(\yxs =a \|\yH)$ determined by the
marginalization~\eqref{noeq:subpop_average} and does \emph{not} have a
maximum-entropy form with the same number of constraints.
%file: test_ME_same_under_scaling.nb

\paragraph{Second example: subnetwork-distribution constraint.}
\label{nosec:maxent_fullsubpop}
Consider another state of knowledge $\yHb$ leading to the following
properties:
\begin{enumerate}%[wide,label=$\yHb$\arabic*.]
\item the average $\yxs$ of a particular
  subnetwork has a probability distribution $q$:
  \begin{equation}
    \label{noeq:constr_subpop_distr}
    \p(\yxs =a \|\yHb) = q(a);
  \end{equation}
\item the probability distribution for the network, $\p(\yX = \yR
  \|\yHb)$, has maximum relative entropy with respect to the uniform
  one, given the constraint above.
\end{enumerate}

\medskip Then the probability distribution for the network given $\yHb$ is
completely determined and satisfies the symmetry
property~\eqref{noeq:homog_math}:
\begin{multline}
  \label{noeq:pop_distr_maxent2}
  \p(\yX= \yR \|\yHb) =
\exp\Biggl[
\sum_{n a=0}^n \yL_{ a}
%\binom{N}{N\yXf}^{-1}
\binom{n}{n a}
\binom{\yvN-n}{\yvN\yRf-n a}\Biggr]
\\
\!\begin{aligned}[b]
&\text{with $\yL_{ a}$ such that}
\\
&\sum_{\yvN A=0}^\yvN 
%\binom{N}{N A}^{-1}
\binom{n}{n a}
\binom{\yvN-n}{\yvN A-n a}\,
\exp\Biggl[
\sum_{n a=0}^n \yL_{ a}
%\binom{N}{N A}^{-1}
\binom{n}{n a}
\binom{\yvN-n}{\yvN A-n a}\Biggr]
=q( a)
\end{aligned}
\end{multline}
(the normalization constraint being unnecessary since $q$ is normalized).
This result is just another application of the maximum-entropy procedure
with $n+1$ (linear) constraints given by \eqn~\eqref{noeq:marginal}, where
the left-hand side is now given and equal to $q(a)$.

This example is equivalent to the generalization of the previous one with
$n$ moment constraints, since knowledge of $\p(\yxs=a \| \yHb)$ is
equivalent to knowledge its first $n$ moments.

***

In this note we would like to analyse and warn about a subtle assumption
behind the \me\ method when it is applied to a population. It can informally
be put this way:
\begin{quote}
  \emph{the \me\ method assumes that the population it is applied to is
    completely isolated from any larger population.}
\end{quote}

\mynote{Version 1}
The \me\ method does not construct a probability distribution out of
nothing, but starting from a uniform distribution. A uniform distribution
is an innocuous assumption for a set of non-composite events, like the
outcomes of a die roll, and also for some sets of composite events, like
the outcomes of the roll of two dice. In the latter case multiplicities
appear.

When applied to a subpopulation, the \me\ method assumes that the uniform over
the larger population is uniform, and therefore factorizable. The new
distribution of the subpopulation will not be uniform, but that of the full
population will still be factorizable into the one for the subpopulation and the
rest.

A uniform distribution, however, is not the right one when we suppose that
learning about an event may tell us something about a related event. For
example, consider 1000 tosses of a particular coin and assume a uniform
distribution over the possible $2^{1000}$ outcomes. If we learn that the
first 999 tosses yielded all \enquote{heads}, the probability calculus
tells us that the probability for the 1000th toss is still 50\%/50\%. It
is a consequence of our choice of a uniform distribution: we have
implicitly declared all tosses to be completely independent, completely
\emph{irrelevant} to one another. This fact is well-known in sampling
theory. A more telling example in fact is that of a presidential election
with two candidates: each citizen will vote for one or the other. We do
survey sampling on a large number of citizens to guess the election's
outcome. If we assumed a uniform distribution over the possible
combinations of choices of all citizens, our sampling would be completely
irrelevant for the choices of the rest of the population.

The latter example has many similarities with that of a neuronal binary
population. When we record the neuronal activity of a sample of neurons from a
brain area, we assume that our measurements can tell us something -- no
matter how vague or imprecise -- about the whole brain area. This means
that we are not assuming a uniform distribution over all possible states of
the area.

\mynote{Version 2}
This may come as a surprise. The method simply requires a number of
exhaustive and mutually exclusive events, and if these are composite events
the final distribution may have a multiplicity factor. When we consider the
$2^{\yvN}$ states of $\yvN$ units we are not excluding that these might be
marginals of $2^{M}$ states of $M$ units. Each one has the same
multiplicity $2^{M-\yvN}$, but this constant multiplicity factor disappears
by normalization. So the method applies just as in the case of $\yvN$ units
only, right?

Right, but

Right, and that is where the problem lies. This way of counting of
multiplicities assumes an underlying

Wrong. In our reasoning we have made subtle assumptions of independence
between the full population and the subpopulation. The problem is that the
counting of multiplicities is not based on simple enumeration, but already
involves probability considerations. Consider three cases with a full
population of two units, $M=2$, of which we consider one unit, $\yvN=1$.
\begin{itemize}
\item First case: all four states are \emph{possible}. The two states of
  the first unit have multiplicity $2$ each. The usual \me\ distribution
  obtains.
\item Second case: only the states with at most one active unit are
  possible. The state $\yxx_1=1$ of the first unit has multiplicity
  $2$, and $\yxx_1=1$ has multiplicity $1$. The \me\ distribution has
  multiplicity factors.
\item Third case: states with at most one active unit are, say, $10^{9}$
  times more probable than the state with no active units. But all four
  states are \emph{possible}. By enumeration this case is like the first:
  multiplicities $(1,1)$. But by common sense it is more similar to the
  second: multiplicities $(2,1)$ for most practical purposes.
\end{itemize}

This simple example shows that the multiplicity inspection that must
precede a \me\ application already involves probability considerations at
the level of the full population. The usual reasoning by enumeration
implicitly assumes a uniform distribution or at least a \emph{factorizable}
distribution.

***If the distribution is factorizable, however, it means that examination of
the subpopulation \emph{cannot give us any insights about the population it is a
  part of}. This is obviously contrary to the reason why we made neuronal observations.

Consider the following ways of proceeding. We:
\begin{enumerate}
\item have a population with $\yvN$ units, $2^{\yvN}$ possible states
\item expect averages of $\yx$ active neurons and $\yxxs$ active pairs
\item use maximum-entropy to choose a probability distribution for the
  states of the $\yvN$ units conforming to our expectations.
\end{enumerate}

We:
\begin{enumerate}
\item have a population with $M$ units, $2^{M}$ possible states
\item expect that any subpopulation of $\yvN$ units has $\yxs$ active
  neurons and $\yxxs$ active pairs
\item use maximum-entropy to choose a probability distribution for the
  states of the $M$ units conforming to our expectations
\item marginalize to find the probability distribution for the states of
  $\yvN$ units.
\end{enumerate}

  \section{Sketched proofs}
  \label{nosec:derivations}

  Variants of the following derivations and combinatorial considerations
  can be found \eg\ in
  \cites[\chaps~I--IV]{whitworth1867_r1965}[\chap~II]{feller1950_r1968}[\chap~3]{jaynes1994_r2003};
  see also \citep{whitworth1897}.

  To derive the joint probability distribution
  \eqref{noeq:joint_plaus_N_homog} from that for the network
  average~\eqref{noeq:pop_average_Q}, consider that if the network total
  is $\yvN\yXf$, then $\yvN\yXf$ out of $\yvN$ units are active, and there are
  $\tbinom{\yvN}{\yvN\yXf}$ possible states for which this can be true; therefore
  \begin{equation}
    \labelbis{eq:joint_plaus_\yvN_homog}
    \p(\yX = \yR \| H) = \binom{\yvN}{\yvN\yRf}^{-1}Q(\yRf).
  \end{equation}
An analogous reasoning for $n$ and $\yxs$ leads to an analogous equality,
\begin{equation}
  \label{noeq:aver_subpop}
  \p(\yx=\yr \| H) = \binom{n}{n\yrs}^{-1} \p(\yxs=\yrs\|\yH),
\end{equation}
for the subnetwork.

\bigskip

Let us next consider the probability $\p(\yxs =a \| \yXf = A,\yH)$ for
the subnetwork average $\yxs$ conditional on the network average $\yXf$.
There are $\tbinom{\yvN}{\yvN\yXf}$ possible network states if the network
average is $\yXf$, \ie\ if $\yvN\yXf$ units are in state $1$; the conditional
probability of each is therefore $1/\tbinom{\yvN}{\yvN\yXf}$, owing to the
symmetry assumption~\eqref{noeq:homog_math}. Now consider the subnetwork of the
first $n$ units. The conditional probability of having $n\yxs$ specific
ones in state $1$ is the sum of the probabilities of all states for
which $\yvN\yXf-n\yxs$ of the remaining $\yvN-n$ units are in state $1$; there are
$\tbinom{\yvN-n}{\yvN\yXf-n\yxs}$ such states, all equally probable. Finally,
there are $\tbinom{n}{n\yxs}$ possible ways, all equally probable, in which
$n\yxs$ of the first $n$ units can be in state $1$. In formulae,
\begin{equation}
\label{noeq:conditional_prob_explained}
\begin{split}
      \p(\yxs =a \| \yXf=A, \yH)
      &=
      \sum_{\yr}^{\yrs=a} \sum_{\yR}^{\yRf=A} 
      \p(\yx =\yr \|  \yX = \yR, \yH) \,
      \p(\yX =\yR  \| \yXf=A,\yH),
      \\
      &=
      \sum_{\yr}^{\yrs=a} \sum_{\yR}^{\yRf=A} 
      \p(\yx=\yr \|  \yX =\yR, \yH) \,
      \binom{\yvN}{\yvN A}^{-1},
      \\
      &=\sum_{\yr}^{\yrs=a} 
      \binom{\yvN-n}{\yvN A-n a}\binom{\yvN}{\yvN A}^{-1},
      \\
      &=
      \binom{n}{n a}\binom{\yvN-n}{\yvN A-n a}\binom{\yvN}{\yvN A}^{-1},
\end{split}
\end{equation}
  which is the conditional probability~\eqref{noeq:conditional_prob}. Note
  that this is just a derivation of the hypergeometric distribution
  \cites[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968},
  which describes the probability of, say, drawing a proportion of $\yxs$
  blue balls in $n$ drawings without replacement from an urn with $\yvN$
  balls, a fraction $\yXf$ of which are blue.

  \bigskip The probability of a subnetwork average $\yxs$ is then, by
  marginalization,
  \begin{equation}
    \label{noeq:subpop_average_explained}
    \begin{split}
      \p(\yxs =a \| \yH)
      &= \sum_{\yvN A=0}^\yvN\p(\yxs = a \| \yXf = A, \yH)\,
      \p(\yXf = A \| \yH),
      \\
      &=\sum_{\yvN A=0}^\yvN
      \binom{n}{n a}\binom{\yvN-n}{\yvN A-n a}\binom{\yvN}{\yvN A}^{-1}
      Q(A).
    \end{split}
  \end{equation}
  which proves the subnetwork-average formula~\eqref{noeq:subpop_average}.
  This formula, combined with \eqns~\eqref{noeq:aver_subpop} and
  \eqref{noeq:joint_plaus_N_homog}, leads to the conditional
  probability~\eqref{noeq:conditional_total_state}.

  \bigskip The independence of the expectation of products of states from
  the subnetwork size is trivial by marginalization:
  \begin{multline}
    \label{noeq:margin_products}
    \sum_{\yx} \yvx_1\dotsm \yvx_m \p(\yx =\yr \| \yH)
      =\sum_{\yr,\yR}  \yvr_1\dotsm \yvr_m \p(\yx=\yr \| \yX=\yR,\yH)
      \, \p(\yX=\yR \| \yH),
      \\
      \begin{aligned}
      &=\sum_{\yr,\yR} \yvr_1\dotsm \yvr_m\,\delt(\yvR_1-\yvr_1)\dotsm\delt(\yvR_m-\yvR_m)
      \p(\yX = \yR \| \yH),
      \\
      &=\sum_{\yR} \yvR_1\dotsm \yvR_m \p(\yX =\yR \| \yH).
    \end{aligned}
  \end{multline}
  All such $m$-fold products have the same expectation by symmetry,
  therefore their subnetwork average will do, too, being an average of
  equal terms.

  Now consider the sum of all distinct products of states of two units in
  the subnetwork:
  \begin{equation*}
    \yvx_1 \yvx_2 + \yvx_1 \yvx_3 + \dotsb + \yvx_{n-1} \yvx_n.
  \end{equation*}
  The terms in this sum are either $0$ or $1$. The non-vanishing ones are
  those with index pairs chosen from the $n\yxs$ units of the subnetwork
  which are in state $1$, and there are $\tbinom{n\yxs}{2}$ such choices,
  so the sum above is equal to $\tbinom{n\yxs}{2}$. The sum has
  $\tbinom{n}{2}$ terms, so their average is
  $\tbinom{n\yxs}{2}/\tbinom{n}{2}$. Generalizing the argument to products
  of $m$ units, we have that
  \begin{equation}
    \sav{\yvx_{i_1} \dotsm \yvx_{i_m}} = \binom{n\yxs}{m} \binom{n}{m}^{-1}.
  \end{equation}
  Then, using \eqn~\eqref{noeq:subpop_average},
  \begin{multline}
    \label{noeq:expect_product_m}
    \expe*{\sav{\yvx_{i_1} \dotsm \yvx_{i_m}} \| \yH}
    \\
    \begin{aligned}[b]
      &=\tfrac{(n-m)!}{n!}
      \expe*{m!\,\tbinom{n\yxs}{m}  \| \yH}
      \\
      &\equiv\frac{(n-m)!}{n!}
      \sum_{n a=0}^n m!\, \binom{n a}{m} \p(\yxs =a \|\yH),
      \\
      &=\frac{(n-m)!}{n!} \sum_{\yvN A=0}^\yvN
      Q( A)
      \Biggl[ 
      \sum_{n a=0}^n 
      m!\,\binom{n a}{m}\, \binom{n}{n a}\binom{\yvN-n}{\yvN A-n a}\binom{\yvN}{\yvN A}^{-1} 
      \Biggr].
    \end{aligned}
  \end{multline}
  The expression in brackets is the $m$th factorial moment of the
  hypergeometric function, and is given by \citep{potts1953}
%file: factorial_moments_etc.nb
  \begin{equation}
    \sum_{n a=0}^n m!\,\binom{n a}{m} \binom{n}{n a}\binom{\yvN-n}{\yvN\yXf-n a}
    \binom{\yvN}{\yvN\yXf}^{-1} 
    =
    m!\,\binom{n}{m} \binom{n a}{m}\binom{\yvN}{m},
  \end{equation}
  which combined with the previous equation yields the second line of
  \eqn~\eqref{noeq:expe_products}; its last equality comes from the identity
  \begin{equation}
    \label{noeq:binom_identity}
    \binom{\yvN}{M}\binom{M}{m}=\binom{\yvN}{m}\binom{\yvN-m}{M-m},
  \end{equation}
  easily derived by writing the binomial coefficients in terms of
  factorials. Finally, \eqns~\eqref{noeq:moments}, relating the moments of
  the distributions for subnetwork and network averages, is obtained
  from the definition of moments,
  \begin{equation}
    \label{noeq:moments_def}
    \expe*{\yxs^m \| \yH} \defd
    \sum_{n a=0}^{n} a^m \p(\yxs = a  \| H),
    \quad
    \expe{\yXf^m \| \yH} \defd
    \sum_{\yvN A=0}^{\yvN} A^m \p(\yXf = A  \| H),
  \end{equation}
  replaced in the equalities for the factorial
  moments~\eqref{noeq:eq_factmom_prod}, by recursively solving in terms of
  the moments of the network distribution.

  *******

Here is an example.

Suppose our desired constraints are $\expe{\yxs } = c_1$ and
$\expe{\yxxs} = c_2$, the expectations being given by the fourth term of
\eqn~\eqref{eq:expe_products}. Applying \me\ to the distribution of the
average of the full population we find
\begin{equation}
  \label{eq:full_pop_av_example_ME}
  \p(\yXf = \yRf \| \yHa)  = \yL\,
\binom{\yvN}{\yvN \yRf}\exp\biggl[
\yL_2 \yRf \,\frac{\yvN \yRf-1}{\yvN-1} + \yL_1\, \yRf\biggr]
\end{equation}
with such parameters $\yL$, $\yL_1$, $\yL_2$ as to satisfy normalization
and constraints. We call this state of knowledge $\yHa$. Marginalization by
\eqn~\eqref{eq:marginal} then gives
\begin{equation}
  \label{eq:full_pop_av_example_ME_marg}
  \p(\yxs = \yrs \| \yHa)  = \yL
  \sum_{\yvN\yRf=0}^\yvN
  \binom{n}{n\yrs} \binom{\yvN-n}{\yvN\yRf-n\yrs} \exp\biggl[
\yL_2 \yRf \,\frac{\yvN \yRf-1}{\yvN-1} + \yL_1\, \yRf\biggr].
\end{equation}
Applying \me\ to the distribution of the average of the sample, using the
third term of \eqn~\eqref{eq:expe_products}, we find
\begin{equation}
  \label{eq:sample_av_example_ME}
  \p(\yxs = \yrs \| \yHb)  = \yl\,
  \binom{n}{n\yrs}
  \exp\biggl[\yl_2 \yrs \,\frac{n \yrs-1}{n-1} + \yl_1\, \yrs\biggr]
\end{equation}
with such parameters $\yl$, $\yl_1$, $\yl_2$ as to satisfy the constraints.
We call this state of knowledge $\yHb$.

The resulting distributions $\p(\yxs = \yrs \| \yHa)$,
$\p(\yxs = \yrs \| \yHb)$ are different.
Figure~\ref{fig:diff_maxent_pop_sample} shows one example, for $\yvN=5000$,
$n=100$, $c_1=0.45$, $c_2=0.35$. The two distributions have slightly
different modes and the one obtained from \me\ at the population level is
more peaked. Another example is shown in
\fig~\ref{fig:diff_maxent_pop_sample_realdata}, for $\yvN=5000$, $n=200$,
and neurobiologically more realistic constraints $c_1=0.045$, $c_2=0.0025$
\cite{rostamietal2016}. The discrepancy is maybe not as large as in the
previous example, but the tails of the distributions are still very
different.

**********************************************************

The neurons recorded in these kinds of experiments are usually picked out
according to an unknown process, and from their observation we expect to
learn something about all other neurons in the same brain area. That is,
they are assumed to be a \enquote{representative random sample} of the
neurons in that area. Our discussion hinges on this often just tacitly
understood assumption.

One of the uses of \me\ methods

number. It is large from large
population of neurons [\cite]. The recorded neurons are usually picked out
from a specific brain area according to an unknown process, and from their
observation in relation to stimuli or behavioral task, we strive for
understanding the mechanism behind neuronal coding\mynote{LM still don't
  know what people mean by this, can we find an unambiguous term or
  rephrase?}. That is, they are assumed to be a \enquote{representative
  random sample} of the neurons in that area.\mynote{This assumption says
  that from the sample we can learn about the population. The relation of
  population activity with stimuli or behaviour is a separate problem, so
  \enquote{this is} doesn't really fit...}

Huge amounts of parallel recordings from hundreds of neurons, available
today, are accompanied by the statistical challenges to be employed for
shedding light on the mechanism behind neuronal coding. The principle of
\me\ was proposed as a way to overcome these challenges by constructing a
full description of the neuronal firing patterns. The \me\ method employs
only first and second order statistics of the representative random sample
to construct a \enquote{maximally noncommittal} \cite{jaynes1963}
probability distribution of the firing patterns [\cite]. \mynote{LM:
  \enquote{noncommittal} doesn't mean anything to me to be honest, even if
  Jaynes used that term. I would like to openly say this, backed up by
  references, as done some paragraphs below.\\It seems they have used \me\
  for other purposes too, sometimes unclear ones. And with all kinds of
  constraints. So we shouldn't be too specific here in relation to purpose
  and constraints. I'd rather give references and the readers can form
  their own ideas about its use.}

Using the pairwise \me\ models many studies investigated the questions such
as, which level of interaction\mynote{ambiguous word} (pairwise or higher-order)
between neurons is suffiecient to explain the observed experimental data
[\cite]? Or how collective behavior of neurons can be related to stimuli or
behavioral tasks [\cite]? add more **

Here we show that there is a quandary in applying \me\ to a representative
random sample of neurons to generate a maximally noncommital distribution
for that sample.

The quandary is this. Assume that our sample of neurons is
representative of some population. If we assign a \me\ distribution to the
sample, then we cannot assign a \me\ distribution to the full population.
Vice versa, If we assign a \me\ distribution to the full population, then
we cannot assign a \me\ distribution to the sample. This impossibility
holds at least for \me\ distribution with moment constraints, and even if
the orders of the moments constrained in the sample and in the full
population are different.

In other words, if our sample is a \enquote{random representative} of some
population, then we must choose which has a \enquote{maximally
  noncommittal} distribution: the sample or the population. We can't choose
both. It goes without saying that if one is \enquote{maximally
  noncommittal}, the other must be somehow \enquote{committal}. Which
choice is most meaningful?

In the rest of the paper we will mathematically show the contradiction
above and generate a \me\ distribution for the full population rather than
the sample. We will also present some probability relations relevant to
random sampling. These relations are well-known in survey sampling and in
the pedagogic problem of drawing from an urn without replacement, yet they
are somewhat hard to find explicitly written in the neuroscientific
literature.

 \mynote{Version 1}
The present note has three nested purposes.

The first and more general is to remind our readers that the \me\ method
rests on subtle assumptions that may be contradictory with other
assumptions we want to make on the data under study.

As a concrete example we show -- and this is the second purpose -- that a
sample of a neuronal population cannot be assigned a \enquote{maximally
  noncommittal} \me\ distribution and at the same time be considered a
\enquote{random representative} of the population.

The proof of the example above rests on some probability relations from
classical random sampling. These relations are somewhat hard to find
explicitly stated in the neuroscientific literature; the third, minor
purpose of this note is to state them explicitly.

Let's clarify at once that our discussion pertains to a binary population at
one specific instant or short window in time. It is possible to
similarly discuss multi-state neuron models and population dynamics; but for
simplicity neurons are here assumed to be in a fixed, active \enquote{$1$}
or inactive \enquote{$0$} state; and evolution, change, time correlations,
and similar concepts do not concern us.

Maximum-entropy models are used for a variety of (sometimes not completely
clear) reasons. For our purposes let's say that the \me\ method is assumed
to generate a \enquote{maximally noncommittal} \cite{jaynes1963} probability
distribution. We intend the quoted expression only as an umbrella term,
also because it means very little without further clarifications.

The neurons recorded in these kinds of experiments are usually picked out
according to an unknown process, and from their observation we expect to
learn something about all other neurons in the same brain area. That is,
they are assumed to be a \enquote{representative random sample} of the
neurons in that area. Our discussion hinges on this often just tacitly
understood assumption.

\mynote{Version 2 }
In this note we would like to show that there is a
contradiction in applying \me\ to a \enquote{representative random sample}
of neurons to generate a \enquote{maximally noncommital} distribution for
that sample.
% \\
% \mynote{V2 }In this note we would like to show that there is a clash between the
% \enquote{representative random sample} spirit and the \enquote{maximally
%   noncommital} spirit of \me\ models applied to such a sample of neurons.

The contradiction is this. Assume that our sample of neurons is
representative of some population. If we assign a \me\ distribution to the
sample, then we cannot assign a \me\ distribution to the full population.
Vice versa, If we assign a \me\ distribution to the full population, then
we cannot assign a \me\ distribution to the sample. This impossibility
holds at least for \me\ distribution with moment constraints, and even if
the orders of the moments constrained in the sample and in the full
population are different.

In other words, if our sample is a \enquote{random representative} of some
population, then we must choose which has a \enquote{maximally
  noncommittal} distribution: the sample or the population. We can't choose
both. It goes without saying that if one is \enquote{maximally
  noncommittal}, the other must be somehow \enquote{committal}. Which
choice is most meaningful?

In the rest of the paper we will mathematically show the contradiction
above and generate a \me\ distribution for the full population rather than
the sample. We will also present some probability relations relevant to
random sampling. These relations are well-known in survey sampling and in
the pedagogic problem of drawing from an urn without replacement, yet they
are somewhat hard to find explicitly written in the neuroscientific
literature.
\fi
\end{document}